# (PART) Preliminaries {-}

# Introduction {#intro}

This chapter provides an overview of the purpose of developing the **medfate** an **medfateland** R packages, their main simulation functions and expected applications. 

## Purpose

Being able to anticipate the impact of global change on forests is one of the major environmental challenges in contemporary societies. However, uncertainties in how forest ecosystems function and practical constraints in how to integrate available information still hinder the availability of robust and reliable predictive models. Despite the amount of knowledge accumulated about the functioning and dynamics of Mediterranean forests, scientists should make coordinate their efforts to address the challenge of integrating the different global change drivers into simulation tools useful for research and applications.

### Package medfate

The R package **medfate** has been designed as a platform to simulate the functioning and dynamics of individual forest stands at temporal scales from days to years. Climate is the main environmental driver covered by the package and the ecological, hydrological and physiological processes surrounding these are fundamental for the simulation models included in the package. In particular, the package allows the simulation of water balance of soils and plants within forest stands. It also allows simulating plant growth of a set of cohorts competing for light and water within a forest stand. Finally, the package functions to relate the amount of plant biomass and the water status of plant tissues to fuel characteristics and, hence, fire hazard.

The version of the reference manual that you are reading is intended to reflect **version `r packageVersion("medfate")`** of the package.

### Package medfateland

The R package **medfateland** has been designed to extend the capabilities of medfate to the landscape level. It allows running the stand-level models available in medfate on points and cells within landscape. More importantly, medfateland allows considering spatial hydrological processes in forested watersheds. Hence, medfateland can be used as a tool for eco-hydrological applications.   

The version of the reference manual that you are reading is intended to reflect **version `r packageVersion("medfateland")`** of the package.

## Package installation

### medfate

Package **medfate** can be found at [CRAN](https://CRAN.R-project.org/package=medfate), where it is updated every few months. Hence, it can be installed using:
```{r eval=FALSE}
install.packages("medfate")
```

Users can also download and install the latest stable versions GitHub as follows (required package `devtools` should be installed/updated first):

```{r installation, eval = FALSE}
devtools::install_github("vegmod/medfate")
```

When installing from GitHub, may need to force the installation of package vignettes, by using: 

```{r installation vignettes, eval = FALSE}
devtools::install_github("vegmod/medfate", 
                         build_opts = c("--no-resave-data", "--no-manual"))
```

Note that some package dependencies are only *suggested*, so that they are not automatically installed when installing **medfate**.

### medfateland

Package **medfateland** is currently only available at GitHub, and can be installed using: 

```{r installation medfateland, eval = FALSE}
devtools::install_github("vegmod/medfateland")
```

### Installation from sources

Both packages have a user interface in R, but actual calculations have been implemented as C++ functions and linked to R via the **Rcpp** package. This means that you will need a compiler to install **medfate** and **medfateland** packages from sources (i.e. Rtools in Windows).

## Data structures and functions in medfate

### Data structures

Package **medfate** deals with simulations at stand level and, hence, the data structures defined in the package are meant to encapsulate information about a target forest stands and the properties of the soil where it grows:

+ S3 class `soil` is a list of soil physical and hydraulic properties, including both parameters and state variables. Soil objects are described in subsection \@ref(soilinitialization).
+ S3 class `forest` is a list of forest inventory data (i.e. mainly tree and shrub measurements) corresponding to a single forest stand. Forest plot objects are described in subsection \@ref(forestobjects). 
+ S3 classes `spwbInput` and `growthInput` are lists containing all the vegetation parameters and state variables necessary to run water-balance simulation functions and forest growth simulation functions (see below). 

### Dynamic simulation functions

Forest water and energy balance can be studied for a given forest stand using function `spwb()` (soil-plant-water-balance). This function allows simulating vertical hydrological processes (rainfall interception, soil infiltration, percolation and evapotranspiration) and plant physiological processes related to plant transpiration (hydraulics, photosynthesis and stomatal regulation).  The same function is used to run two models of different degree of complexity, from which the user should choose depending on the intended application and data availability. Most processes are implemented at the daily scale, although some operate at subdaily time steps. Function `spwb()` has a closely related function `spwb.day()` that allows focusing on subdaily processes. Finally, one can simulate plant physiological processes only, while having soil moisture as a dynamic input like weather, using function `pwb()` (plant-water-balance).

Changes in primary (leaf area) and secondary (wood) growth are key to evaluate the influence of climatic conditions on plant and forest structure and function. Function `growth()` extends the previous models because it allows simulating growth and mortality of a set of plant cohorts competing for light and water in a given forest stand. It can be used to forecast temporal variation in water fluxes and soil water content in forest stands while taking into account plant acclimation. Analogously to the simulation of water balance, `growth()` has a closely related function `growth.day()` that allows focusing on subdaily processes (i.e. carbon balance). Importantly, `growth()` function and the processes that it implements are still **under development**.

### Sub-model functions

Many of the sub-models included in **medfate** are implemented as C++ functions and internally called by the simulation functions implementing top-level models. Most sub-models have, however, their corresponding R function to made them directly available to the user. This facilitates understanding the different sub-models and a more creative use of the package. Sub-model functions are grouped by *subject*, which is included in the name of the function. The different sub-model functions are (by subject):

* `biophysics_*`: Physical and biophysical utility functions.
* `carbon_*`: Structural and non-structural carbon compartments and flows.
* `hydraulics_*`: Plant hydraulics.
* `hydrology_*`: Canopy and soil hydrology.
* `light_*`: Light extinction and absortion.
* `moisture_*`: Tissue water relations and live tissue moisture.
* `pheno_*`: Leaf phenology.
* `photo_*`: Leaf photosynthesis.
* `root_*`: Root distribution and root/rhizosphere conductance estimation.
* `soil_*`: Soil hydraulics and thermodynamics.
* `spwb_*`: Soil water balance routines.
* `transp_*`: Stomatal regulation and plant transpiration.

### Static functions

Package **medfate** includes a number of functions to examine static properties of plants conforming forest stands, summary functions at the stand level or vertical profiles of several physical properties:

* `plant_*`: Cohort-level information (species name, id, leaf area, height...).
* `species_*`: Cohort-level attributes aggregated by species (e.g. basal area).
* `stand_*`: Stand-level attributes (e.g. basal area).
* `vprofile_*`: Vertical profiles (light, wind, fuel density, leaf area density).

Vegetation functioning and dynamics have strong, but complex, effects on fire hazard. On one hand, growth and death of organs and individuals changes the amount of standing live and dead fuels, as well as downed dead fuels. On the other, day-to-day changes in soil and plant water content changes the physical properties of fuel, notably fuel moisture content. Package **medfate** provides functions to estimate fuel properties and potential fire behaviour in forest inventory plots. Specifically, function `fuel_stratification()` provides a stratification of the stand into understory and canopy strata; and `fuel_FCCS()` calculates fuel characteristics. A fire behaviour model is implemented in function `fire_FCCS()` to calculate the intensity of surface fire reaction and the rate of fire spread of surface fires assuming a steady-state fire. Fuel and fire behaviour functions allow obtaining the following: 

1. Fuel characteristics by stratum.
2. Surface fire behavior (i.e. reaction intensity, rate of spread, fireline intensity and flame length).
3. Crown fire behavior.
4. Fire potential ratings of surface fire behavior and crown fire behavior.

## Data structures and functions in medfateland

### Data structures

While simulation functions in the package are run on individual forest stands, the package has been designed so that it could be part of a broader modelling infrastructure including *spatial processes* such as lateral water flows.

### Dynamic simulation functions
