# Model inputs

Data input for simulation models can be grouped into four categories. First, a description of the  **vegetation** existing in the target forest stand is needed. This information should be available at a level of detail appropriate for the model design, which in our case includes not only species composition and/or structure, but also functional parameters of woody plants. Vegetation structure is considered static in `spwb()` calls, but becomes dynamic in calls to `growth()`. Hence, in this latter case the input vegetation describes to the starting point. Second, **soil** information is also required, because underground water resources are important for the processes included in the model. While soil features are considered static parameters in model simulations, **weather** is always a dynamic input that defines the temporal variation of environmental conditions in which physical and biological processes occur. Finally, **topographic and geographic information** is often required, since internal calculations of some physical variables (such as atmospheric pressure or radiation) require them. 

This chapter describes the requirements of **medfate** package with respect to each of these data input categories. The last section describes additional parameters that are used to specify global options of simulation functions (i.e., `spwb()`, `pwb()`, `growth()`, `fordyn()` and daily simulation functions).

Since preparing inputs for simulations can be hard, package [**medfateutils**](https://github.com/emf-creaf/medfateutils) provides a set of utility functions aimed at facilitating this task.

## Latitude and topography {#topoinput}

Topography has a relatively small direct influence in **medfate** simulations, because most topographic effects are assumed to be taken into account in the estimation of weather input. Nonetheless, *elevation* in meters is needed to estimate atmospheric pressure and air density (see \@ref(meteoinput)), whereas *aspect* (in degrees from North) and *slope* (also in degrees) are needed in addition to *latitude* (also in degrees North) to estimate sunrise/sunset hours and potential radiation. Topographic variables should be readily accessible for users having access to a digital elevation model.

## Soil description  {#soilinput}

Soils are described in **medfate** using $S > 1$ different soil layers ($s \in \{1, 2, ..., S\}$). Here the word *soil* refers to depths that plant rooting systems can reach (i.e., the critical zone), including cracks within the bedrock. The number and size of soil layers may reflect changes in soil properties along depth, but also can be chosen to reflect differences in plant rooting depths. The physical properties of the soil are needed to estimate soil hydraulic properties. Most soil properties are considered *static* parameters in simulations, except soil moisture content, variables that depend on moisture (e.g. soil water potential and soil conductivity) and soil temperature, if the soil energy balance is considered in simulations.

### Soil physical properties

Soil physical properties, such as *texture* (i.e. volume percent of sand, silt and clay), *organic matter content*, *bulk density* or *rock fragment content*, can differ between soil layers, and this has important consequences for soil water retention capacity and soil hydraulics. Specifying layers with an elevated rock fragment content (e.g. layers of weathered rock where soil particles are scarce) may be important in  seasonally-arid climates like the Mediterranean, because plants often extend their roots into cracks existing in the parent rock to access water during summer [@Ruffault2013]. 

For each soil layer $s$, the following physical parameters are needed:

| Symbol | Units | R  | Description                         | 
|--------|-------|----| ------------------------------------|
| $d_{s}$ | mm | `widths` | Width of soil layer $s$ |
| $P_{clay,s}$ | % | `clay` | Percent volume of clay (within soil particles fraction) in layer $s$  |
| $P_{sand,s}$ | % | `sand` | Percent volume of sand (within soil particles fraction) in layer $s$ |
| $OM_s$ | % | `om` | Percentage of organic mater per dry weight (can be missing) in layer $s$ |
| $BD_{s}$ | $g \cdot cm^{-3}$ |`bd` | Bulk density in layer $s$|
| $P_{rocks,s}$ | % | `rfc` | Rock fragment content as percent volume (within whole soil) in layer $s$ |

The percent volume of silt is 100\% minus the percent volume of clay and sand. Whenever possible, soil physical properties should be measured in soil profiles conducted at the target forest plot (although soil profiles rarely reach rooting depths of deeply-rooted plants). Soil input data should be arranged in a `data.frame` with soil layers in rows and physical variables in columns (see function `defaultSoilParams()`). 

Package [**medfateutils**](https://github.com/emf-creaf/medfateutils) includes a function `soilgridsParams()` to fetch soil information from [SoilGrids.org](https://soilgrids.org/), a global soil database currently providing soil data at 250m scale. This can be helpful to users lacking local soil measurements, but the uncertainty of SoilGrids estimates can be high for some areas and soil properties, especially soil depth and rock fragment content.

The depth of a given soil layer $s$ ($Z_s$; in mm) is defined as the sum of layer widths from the surface to the target layer:
\begin{equation}
Z_s = \sum_{i=1}^s{d_s}
\end{equation}
And the overall *soil depth*  ($Z_{soil}$; in mm) is the sum of soil layer widths. As it may include rocky layers, the value of $Z_{soil}$ may be larger than the soil depth reported in soil profiles. 

### Water retention curves {#waterretentioncurves}

The water retention curve of a soil is the relationship between *volumetric soil moisture content* ($\theta$, in $m^3 \cdot m^{-3}$ of soil, excluding rock fragments) and the corresponding *soil water potential* ($\Psi$, in MPa), i.e. the amount of work that must be done per unit quantity of pure water in order to transport reversibly and isothermally an infinitesimal quantity of water from a reference pool. The shape of the water retention curve (also called the *soil moisture characteristic curve*) depends on physical properties (mainly texture and bulk density, but also organic matter content). Since soil layers usually differ in their physical properties they also normally have different water retention curves.

Two water retention curve models are available in **medfate** (fig. \@ref(fig:retention-curves)):

1. *Saxton model*: In this model, volumetric soil moisture $\theta(\Psi)$ corresponding to a given water potential $\Psi$ (in MPa) below field capacity (i.e., $\Psi < \Psi_{fc}$) is calculated using: 
\begin{equation}\theta(\Psi) = (\Psi/A)^{(1/B)}\end{equation}
where $A$ and $B$ depend on the texture and, if available, organic matter in the soil. If organic matter is available, $A$ and $B$ are calculated from $P_{clay}$, $P_{sand}$ and $OM$ following @Saxton2006. Otherwise, they are calculated from $P_{clay}$ and $P_{sand}$ as indicated in @Saxton1986. Volumetric changes between field capacity and saturation (i.e., $\Psi_{fc} \leq \Psi < 0$) are estimated using a linear function.
2. *Van Genuchten model*: The well known van Genuchten [-@Genuchten1980] model is:
\begin{equation}\theta(\Psi) = \theta_{res}+\frac{\theta_{sat}-\theta_{res}}{\left[1+ (\alpha \cdot \Psi)^n \right]^{1-1/n}}\end{equation}
where $\theta(\psi)$ is the water retention, $\theta_{sat}$ is the saturated water content, $\theta_{res}$ is the residual water content, $\alpha$ is related to the inverse of the air entry pressure, and $n$ is a measure of the pore-size distribution. 

(ref:retention-curves-cap) Water retention curves under the Saxton and Van Genuchten models, for the same physical attributes (25% sand, 25% clay, 50% silt and bulk density of $1.5 g \cdot cm^{-3}$). Parameters of the Van Genuchten curve have been estimated using pedotransfer functions given in @Toth2015.

```{r retention-curves, fig.align='center', fig.width=4, fig.height=3.5, echo = FALSE, fig.cap='(ref:retention-curves-cap)'}
soil_retentionCurvePlot(soil(defaultSoilParams()), model="both")+
  theme(legend.position =c(0.7,0.8))
```


### Soil initialization  {#soilinitialization}
Soil initialization is needed to estimate soil hydrological parameters for each soil layer $s$ from their physical attributes. Initialization is done using function `soil()`, which returns an object of a S3 class with the same name. Simulation functions in **medfate** need an object of class `soil` to be run. 

Function `soil()` adds the following information to the physical soil description: 

| Symbol | Units | R  | Description                               | 
|--------|-------|----| ------------------------------------------|
| $P_{macro, s}$ | % | `macro` | Percentage of macroporosity corresponding to each soil layers | 
| $\gamma_{soil}$ | $mm \cdot day^{-1}$ | `Gsoil` | The maximum daily evaporation rate from soil (see \@ref(soilevaporation)) |
| $\kappa_{soil}$ |  | `Ksoil` | Extinction parameter to regulate the amount of water extracted from each soil layer when simulating evaporation from bare soil (see \@ref(soilevaporation)) |
| $\theta_{sat, s}$ |$m^3 \cdot m^{-3}$ | `VG_theta_sat`| Volumetric soil moisture at saturation |
| $\theta_{res, s}$ |$m^3 \cdot m^{-3}$ | `VG_theta_res` | Residual volumetric soil moisture |
| $n_s$ | | `VG_n` | Parameter $n$ of the Van Genuchten [-@Genuchten1980] model |
| $\alpha_s$ | $MPa^{-1}$ | `VG_alpha`| Parameter $\alpha$ of the Van Genuchten [-@Genuchten1980] model |
| $K_{sat,s}$ | $mmol \cdot s^{-1} \cdot m^{-1} \cdot MPa^{-1}$ | `Ksat` | Saturated soil conductivity |
| $K_{drain}$ | $mm \cdot d^{-1}$ | `Kdrain` | Hydraulic conductivity for deep drainage. |

Macroporosity values are calculated using the equations given in @Stolf2011, while $\gamma_{soil}$ and $\kappa_{soil}$ are derived from texture. Parameters of the van Genuchten model for each layer are estimated from the physical description of the layer using one of two pedotransfer functions (see help for `soil()`): 

1. Using the USDA texture classification and the average texture class parameters given by @Carsel1988.
2. Directly from the soil texture, organic matter and bulk density, using the pedotransfer functions in @Toth2015.

Soil layer saturated conductivity ($K_{sat,s}$) is estimated using @Saxton2006 or @Saxton1986, depending on whether an estimate of organic mater is available. $K_{drain}$ is currently initialized from soil saturated conductivity in the deepest soil layer, but should be defined depending on the permeability of the bedrock material.

Besides defining soil parameters, function `soil()` also initializes *soil state variables*: soil moisture content, soil temperature (although set to a missing value) and water table depth.  Users can edit the `soil` object manually (it is actually a `list`), for example to provide specific parameters of the Van Genuchten retention curve calibrated from soil measurements. 

### Water content and water table depth

The *volume of water* ($V_s$ in $mm = l \cdot m^{-2}$ of ground area) in a soil layer $s$ is calculated from its water potential ($\Psi_s$) and water retention curve using:
\begin{equation}
V_s(\Psi_s) = d_s\cdot ((100-P_{rocks,s})/100)\cdot\theta_{s}(\Psi_s)
\end{equation}
where $d_s$ is the depth of the soil layer (in mm) and $P_{rocks,s}$ is the percentage of rock fragments. The overall volume of water in the soil (accessible by roots) is simply the sum of water content over soil layers:
\begin{equation}
V_{soil} = \sum_{s=1}^S{V_s(\Psi_s)}
\end{equation}

A number of fixed water volumes (and their corresponding moisture) are important to remember: 

  + *Water holding capacity* ($V_{fc, s}$, in mm) of the soil layer $s$ is defined as the volume of water at *field capacity*, i.e. at $\Psi_{fc} = -0.033$ MPa, the amount of water held in the soil after gravitational water has drained away:
\begin{equation}
V_{fc,s} = V_{s}(-0.033) = d_s\cdot ((100-P_{rocks,s})/100)\cdot\theta_{fc,s}
\end{equation}
where $\theta_{fc,s} = \theta_s(-0.033)$ is the corresponding moisture at field capacity. 
  + *Water content at saturation* ($V_{sat, s}$, in mm) is calculated anagolously, but replacing $\theta_{fc,s}$ by $\theta_{sat, s} = \theta_s(0)$, the moisture at *saturation*, the amount of soil moisture when all easily drained spaces between soil particles (i.e. macropores) are also filled with water. 
  + *Water content at wilting point* ($V_{wp, s}$, in mm) is calculated replacing $\theta_{fc,s}$ by $\theta_{wp,s} = \theta_s(-1.5)$, the moisture at *wilting point*, i.e. at $\Psi_{fc} = -1.5$ MPa, a conventional amount of soil moisture beyond which plants are assumed have problems extracting water and start to wilt. However, plants can often extract water from much drier soils.
  + The *amount of extractable water* (in mm) is the difference between water content at field capacity and the water content at a conventional minimum water potential, which can be set at wilting point or lower values (by default at -5 MPa in **medfate**).

The depth of water table is the depth of saturated soil (in mm). In medfate, *water table depth* ($Z_{wt}$, in mm) equals soil depth (i.e. $Z_{soil}$) when all soil layers are below field capacity. When some layers are between field capacity and saturation, water table depth is calculated as:
\begin{equation}
Z_{wt} = \sum_{s}{d_s \cdot \min\left[1,(\theta_{sat,s} - \theta(\Psi_s))/(\theta_{sat,s}-\theta_{fc,s})\right]}
\end{equation}


## Vegetation description {#vegetationinput}

The representation of plants within a target plot is done in **medfate** by means of **cohorts** of woody plants (i.e. trees and shrubs). A plant cohort respresents a set of plants that belong to the same species (or functional group) and are more or less of the same *size*. For example, trees in a stand may be grouped by species and diameter class. This representation is chosen so that package functions can be easily applied to **forest inventory data**, for example from national forest surveys. The description of plant cohorts includes a measure of *abundance* of the cohort in the stand, which in the case of trees is *density* as stems per hectare and for shrubs is *percent cover* (shrubs are often multi-stemmed). The difference in measurement units for abundance is the main motivation for distinguishing between tree cohorts and shrub cohorts, rather than their size or growth form. 

Note that one limitation of the representation of vegetation in **medfate** is that it is not spatially-explicit (i.e. plants cannot have explicit coordinates within forest stands) and hence spatial (horizontal) interactions between plant cannot be taken into account explicitly (but see \@ref(plantwaterpools)).  

### Structural attributes of plant cohorts 

A plant cohort $i$ is either a *tree cohort* or a *shrub cohort*, and is defined using a set of structural attributes shown in the following table, where columns **spwb**, **growth** and **fordyn** indicate whether attributes are required by simulation functions `spwb()/pwb()`, `growth()` and `fordyn()`, respectively:

| Symbol | Units | R  | Description                        | spwb | growth | fordyn |
|--------|-------|----| ------------------------------------|-----|-----|-----|
| $SP_i$  |  | `Species` | Species identity |  Y  |  Y  |  Y  |
| $H_i$  | $cm$  | `Height` | Average tree or shrub height |  Y  |  Y  |  Y  |
| $CR_i$ |       | `CR`| Crown ratio (i.e. ratio between crown length and plant height) | Y | Y |  N  |
| $N_i$ | $ind · ha^{-1}$ | `N` | Density of tree individuals | N | Y |  Y  |
| $DBH_i$ | $cm$ | `DBH` | Tree diameter at breast height | N | Y |  Y  |
| $Cover_i$ | % | `Cover` | Shrub percent cover | N | Y |  Y  |
| $LAI^{live}_i$ | $m^2 \cdot m^{-2}$ | `LAI_live` | (Maximum) leaf area index | Y | Y | N |
| $LAI^{dead}_i$ | $m^2 \cdot m^{-2}$ | `LAI_dead` | Dead leaf area index | Y | Y |  N  |
| $LAI^{\phi}_i$ | $m^2 \cdot m^{-2}$ | `LAI_expanded` | Current expanded leaf area index | Y | Y |  N  |
| $Z_{50,i}$ | mm | `Z50` | Depth above which 50% of the fine root mass is located | Y | Y | Y |
| $Z_{95,i}$ | mm | `Z95` | Depth above which 95% of the fine root mass is located | Y | Y | Y |

Height ($H$) values refer to average height of individuals included in the cohort, and the same for crown ratio ($CR$) and diameter at breast height ($DBH$). While plant size (i.e. height or diameter) is relatively easy to tally, other measurements are not usually made in the field. Package **medfate** includes utility functions that provide estimates $CR$ and $LAI$ from forest inventory data (e.g. heights, DBH and density for measured trees), using allometric relationships calibrated for Catalonia (see chapter \@ref(allometricmodels)). 

**Leaf area index**

$LAI$ variables refer to *one-sided leaf area* of plants in the cohort per surface area of the stand. Leaves standing on branches can be alive or dead. We will call *total leaf area index* of a plant cohort $i$ ($LAI^{all}_{i}$) to the sum of dead and live (and unfolded) leaves of the cohort:
\begin{equation}
LAI^{all}_{i} = LAI^{\phi}_{i}+LAI^{dead}_{i}
\end{equation}

For simulations not involving growth, winter deciduous phenology is still simulated, and the current level of leaves unfolded may be lower than the maximum (live) amount of leaves. If there are $c$ plant cohorts, the leaf area index of the whole stand can be calculated for living leaves, unfolded leaves, dead leaves and total leaf area as:
\begin{eqnarray}
LAI^{live}_{stand} &=& \sum_{i=1}^c{LAI_{i}^{live}} \\
LAI^{\phi}_{stand} &=& \sum_{i=1}^c{LAI_{i}^{\phi}} \\
LAI^{dead}_{stand} &=& \sum_{i=1}^c{LAI_{i}^{dead}} \\
LAI^{all}_{stand} &=& \sum_{i=1}^c{LAI_{i}^{all}} = LAI^{\phi}_{stand}+LAI^{dead}_{stand}
\end{eqnarray}

### Forest plot objects {#forestobjects}

While the previous section described the structural forest variables required to run simulations in **medfate**, forest plot data input in the package is easier if one uses a format that follows closely the Spanish national forest inventory. In this format, each forest plot is represented in an object of class `forest`, a list that contains several elements:
```{r}
data(exampleforestMED)
exampleforestMED
```

Among them, the most important items are two data frames, `treeData` (for trees) and `shrubData` for shrubs. Information about herbs is currently only used for fire hazard calculations. Elements `ID` and `patchsize` are only relevant when using medfate functions in a spatial context.

While not strictly necessary (except for `fordyn()`), `forest` objects are a convenient format to start calculations with **medfate**, because there are many static functions that take forest objects as input and calculate leaf area index for each plant cohort. Another pathway to obtain LAI estimations is by specifying foliar biomass values as additional columns of forest objects. 

With the aim to help users in the task of constructing `forest` objects, package [**medfateutils**](https://github.com/emf-creaf/medfateutils) includes functions to map user data in into tables `treeData` and `shrubData`.


### Vertical leaf distribution {#leafdistribution}

The vegetation inputs for height, leaf area index and crown ratio allow the package to determine the vertical distribution of leaves in the stand. The leaf area of any plant cohort is assumed to be distributed vertically following a **truncated Gaussian function** whose standardized values -1.5 and 1.5 correspond to crown base height ($H_{crown,i}$; in $cm$) and total plant height ($H_i$), respectively. Crown base height is defined as the height corresponding to the first living branch. It is calculated from the crown ratio of the cohort ($CR_i$; a proportion between 0 and 1), which in turn can be estimated as explained in \@ref(eq:CrownHeight). 

Simulation models in **medfate** divide the vertical dimension into vertical layers (by default are 100 cm width, but see control parameter `verticalLayerSize`). Let us define $LAI_{i,j}^{all} = LAI_{i,j}^{\phi}+LAI_{i,j}^{dead}$ as the leaf area index of cohort $i$ in layer $j$ including both functional leaves and dead leaves standing on branches. The truncated Gaussian distribution defines the $LAI_{i,j}^{\phi}$ and $LAI_{i,j}^{dead}$ for all plant cohorts and vertical layers.

Dividing the leaf area of a given layer by its width, one obtains the *leaf area density* ($LAD$ in $m^2 \cdot m^{-3}$). Figure \@ref(fig:leafareadensity) illustrates the leaf area density profile (see function `vprofile_leafAreaDensity()`) corresponding to the forest stand described in the example `forest` object that we showed above. 

```{r leafareadensity, fig.width=8, fig.height=4, fig.align='center', echo=FALSE, fig.cap="Leaf area density distribution in a forest stand. The left panel shows the distribution for each plant cohort separately, whereas the right panel shows the overall density distribution."}
g1<-vprofile_leafAreaDensity(exampleforestMED, SpParamsMED, z = seq(0,1000, by=1), byCohorts = T)
g1 <- g1 + theme(legend.background = element_blank(), legend.position = c(0.8,0.8))
g2<-vprofile_leafAreaDensity(exampleforestMED, SpParamsMED, z = seq(0,1000, by=1))
plot_grid(g1, g2, nrow=1, ncol=2)
```

The leaf area density profile determines the ligth extinction rates through the forest canopy. The same truncated Gaussian distribution is used to distribute leaf and small branch biomass along the vertical dimension.

### Vertical root distribution {#rootdistribution}
Roots can be distinguished functionally into two different organs: **coarse roots** are responsible for mechanical support and transport of water and nutrients towards the stem, whereas **fine roots** are responsible for absorbing water and nutrients. 

Analogously to the aboveground vegetation parameters, inputs describing depths corresponding to 50\% and 95\% of fine roots ($Z_{50,i}$ and $Z_{95,i}$) are used to completely specify the distribution of roots across soil layers. The root system of each plant cohort $i$ is described using $FRP_{i,s}$, the proportion of fine roots (with respect to the whole plant) in each soil layer $s$. $FRP_{i,s}$ values are normally defined using the linear dose response model [@Schenk2002; @Collins2007]: 
\begin{eqnarray}
FRP_{i,s} &=& Y_i(Z_s) - Y_i(Z_{s-1}) \\
Y_i(z) &=& \frac{1}{1+(z/Z_{50,i})^{c_i}}
\end{eqnarray}
where $Y_i(z)$ is the cumulative fraction of fine root mass located between surface and depth $z$; $Z_{50,i}$ is the depth (in mm) above which 50\% of the root mass is located; and $c_i$ is a shape parameter related to $Z_{50,i}$ and $Z_{95,i}$ as $c_i  = 2.94 / \ln(Z_{50,i} / Z_{95,i})$. 

The linear dose response model is quite flexible with respect to fine root distribution (see function `root_ldrDistribution()`). The following figure shows the fine root distribution profile of the same cohorts as in fig. \@ref(fig:leafareadensity) (see `vprofile_rootDistribution()`):

```{r root-distribution, echo = FALSE, fig.width=5, fig.height=4, fig.align = 'center', fig.cap = "Examples of root density profile according to the linear dose response model."}
vprofile_rootDistribution(exampleforestMED, SpParamsMED)
```


The actual input to simulation functions `spwb()`, `growth()` and alike, describes fine root distribution using the matrix of $FRP_{i,s}$ values - not $Z_{50,i}$ and $Z_{95,i}$ -, which means that the proportions of fine roots can be modified manually or be defined by another model. In fact, **medfate** also allows calculating $FRP_{i,s}$ values assuming a conic distribution of fine roots (see `root_conicDistribution()`). 

### Plant water pools and root horizontal distribution {#plantwaterpools}

By default, models in **medfate** assume that soil moisture under all plant cohorts is the same (i.e. water sources corresponding to vertical soil layers are shared among cohorts). However, variations in soil moisture beneath plant cohorts (and, implicitly, horizontal variation of soil moisture) can also be simulated if required by the user. Considering plant **water pools** involves partitioning the stand area into fractions corresponding to the abundance of each plant cohort. More specifically, the model defines as many plant water pools as plant cohorts, with proportions defined by their LAI values:
\begin{equation}
f_{pool,i} = \frac{LAI^{live}_i}{LAI^{live}_{stand}}
\end{equation}
Models assume that the rhizosphere of each plant cohort occupies its own water pool but may extend into the water pools under other plant cohorts. In other words, the root systems of different cohorts may overlap horizontally. Moreover, the horizontal overal of root systems will vary across soil layers. A given plant cohort $i$ will have its roots in layer $s$ partitioned among different water pools. We thus need to define $fr_{i,s,j}$, the (horizontal) proportion of fine roots of cohort $i$ in layer $s$ of the water pool $j$, with the restriction that: 
\begin{equation}
\sum_{j}{fr_{i,s,j}} = 1 \,\, \forall i,s
\end{equation}
It is important to realize that proper estimation of $fr_{i,s,j}$ is challenging when we do not have explicit plant coordinates, root lateral widths, etc. At this point, let us assume we have a reasonable estimate of $Vol_{i}$, the **soil volume explored by coarse roots** of an individual of cohort $i$. Assuming that the proportion of fine roots in each layer is proportional to the proportion of total soil volume explored by coarse roots that corresponds to the same layer we have that $Vol_{i,s}$, the soil volume explored by coarse roots in layer $s$ is:
\begin{equation}
Vol_{i,s} = FRP_{i,s}\cdot Vol_{i}
\end{equation}
In **medfate**, coarse roots are represented by *axial* and *radial* components. The length of axial component of a given layer ($L_{axial, i, s}$) is simply equivalent to its bottom depth, whereas the radial component ($L_{radial, i, s}$) is unknown. However, we can assume that the soil volume explored by coarse roots corresponds to a cylinder:  
\begin{equation}
Vol_{i,s} = \pi \cdot L_{radial,i,s}^2 \cdot L_{axial, i, s}
\end{equation}
and substituting the soil volume $Vol_{i,s}$ in the equation above we can isolate $L_{radial,i,s}$ (see function `root_coarseRootLengthsFromVolume()`). Once we have the radial component we can estimate the area ($m^2$) covered by coarse roots of an individual of the cohort in a given layer $s$:
\begin{equation}
Area_{i,s} = \pi \cdot L_{radial,i,s}^2
\end{equation}
We can then compare $Area_{i,s}$ to the area of the water pool for an individual of cohort $i$, which in $m^2$ is:
\begin{equation}
AreaPool_{i} = 10000 \cdot f_{pool,i}/N_{i}
\end{equation}
If $Area_{i,s} \leq AreaPool_{i}$, then the roots of the individual do not exceed the area of the pool, in other words $fr_{i,s,i} = 1$ and $fr_{i,s,j} = 0$ for all $j \neq i$. If
$Area_{i,s} > AreaPool_{i}$ then the excess should be counted as overlap with the other pools. Assuming that plants are randomly distributed (i.e. no clumping) the probability of finding the pool of a cohort is equal to the proportion of its water pool (see function `root_horizontalProportions()`). All these calculations were done assuming $Vol_{i}$ were known. Estimation of $Vol_{i}$ depends on the complexity of the water balance model and is explained in \@ref(horizontalrootoverlap).


### Functional traits and vegetation initialization {#functionaltraits}

Besides the physical representation of vegetation, forest ecosystem models require information regarding vegetation functioning, because this influences the outcome of biophysical, physiological and hydrological processes. In the following chapters, we will indicate for each simulated process its required functional parameters. Normally, functional traits are described at the species level because infra-specific parameters are hard to get. Even at the species level, many functional (e.g. physiological) traits are hard to obtain, so in **medfate** we provide a default species parameter table for woody species found in Catalonia, with many of them occurring elsewhere in the Mediterranean Basin (`SpParamsMED`). Recently, a new parameter data set is being developed for plants in California (`SpParamsUS`). User can modify these data frames to account for intra-specific trait variation (see function `modifySpParams()`). 

The package provides functions `spwbInput()` and `growthInput()` that prepare the vegetation input for simulation functions `spwb()` and `growth()`, respectively. This initialization mostly consists in compiling the necessary functional traits from the species parameter table, following the $SP_i$ attribute of plant cohorts. Some new model parameters are estimated from the initial ones, while taking into account the structural attributes of cohorts (an example of these derived quantities is stem conductance, which estimated from stem xylem conductivity, Huber value and plant height). Users can take the output of these initialization functions and replace parameter values for specific plant cohorts (see function `modifyCohortParams()`). 

### Update of structural variables during simulations

Vegetation characteristics stay constant during simulations using functions `spwb()` or `pwb()`, although the actual (unfolded) leaf area and dead leaf area may vary depending on leaf phenology.

In contrast, growth simulation requires updating the structure of vegetation, i.e. heights, tree diameters and shrub cover. Function `growth()` can modify any of the vegetation attributes. Finally, function `fordyn()` modifies not only structural variables of the initial cohorts but also involves the recruitment of new woody cohorts during simulations.


## Metereological input {#meteoinput}

Weather input data must include variables calculated at the **daily** scale. Weather data should be arranged in a data frame with days in rows and variables in columns. The following table indicates the symbols, units, definitions and the variable name in R.

| Symbol | Units | R param | Description                        | Basic | Advanced |
|--------|-------|---------|------------------------------------|--------|----------|
| $T_{mean}$ | $^{\circ} \mathrm{C}$ | `MeanTemperature` | Mean temperature | Y | Y |
| $T_{min}$ | $^{\circ} \mathrm{C}$ | `MinTemperature` | Minimum temperature | N | Y |
| $T_{max}$ | $^{\circ} \mathrm{C}$ | `MaxTemperature` | Maximum temperature | N | Y |
| $RH_{min}$ | % | `MinRelativeHumidity` | Minimum relative humidity | N | Y |
| $RH_{max}$ | % | `MaxRelativeHumidity` | Maximum relative humidity | N | Y |
| $P$ | $L \cdot m^{-2} = mm$ | `Precipitation` | Precipitation (either rainfall or snow). | Y | Y |
| $PET$ | $L \cdot m^{-2} = mm$ | `PET` | Potential evapotranspiration, preferably calculated using Penman's (1948) equation | Y | N |
| $Rad$ | $MJ \cdot m^{-2}$ | `Radiation` | Solar radiation after accounting for clouds | (opt.) | Y |
| $u$ | $m \cdot s^{-1}$ | `WindSpeed` | Wind speed | Y | Y |
| $C_{atm}$ | $ppm$ | `CO2` | Atmospheric (above-canopy) $CO_2$ concentration | N | Y |

Dates are needed in **medfate** to estimate parameters like solar declination or the day of the year ($DOY$). Importantly, *dates* should be included as row names (in format *year-month-day*) of the input meteorology data frame. Since it allows producing data frames with appropiate variable units and row/column names, we recommend meteorological input to be generated using package **meteoland** [@DeCaceres2018], but many other sources are possible.

Not all simulation models require all weather variables (see rightmost columns in the previous table). In particular, basic water balance modelling (chapter \@ref(basicwaterbalance)) requires much less information than advance water balance modelling (chapter \@ref(advancedwaterbalance)). The weather input requirements for modelling growth (chapter \@ref(growthmodelling)) or forest dynamics (chapter \@ref(dynamicsmodelling)) depend on the complexity of the water balance model used within it.

Whereas the weather variables listed in the table above are input requirements, **medfate** derives other atmospheric variables from them, using also topographic information (see [utility functions](http://emf-creaf.github.io/meteolandbook/miscellaneous.html#physicalutils) of the **meteoland** reference manual):

| Symbol | Units | R param | Description                        |
|--------|-------|---------|------------------------------------|
| $P_{atm}$ | $kPa$ |  | Atmospheric pressure |
| $\rho_{air}$ | $kg \cdot m^{-3}$ |  | Air density |
| $e_{atm}$ | $kPa$ | | Atmospheric water vapor pressure |


## Simulation control {#controlparameters}

Simulation control parameters are a list of global parameter values, initialized using function `defaultControl()`, that the user can modify to change the general behavior of simulation functions. Here is the set of global control parameters currently accepted in **medfate**:
```{r}
names(defaultControl())
```


Control parameters are employed when initializing vegetation inputs and state variables using functions `spwbInput()` and `growthInput()`, since they are needed also to make decisions at the point of initialization. Control parameters are also stored in the result of these functions, so the user does not need to specify again control parameters when calling simulation functions such as `spwb()` or `growth()`. Function `fordyn()` is different than the other simulation functions, because the list control parameters is directly passed to the function.

Not all control parameters are relevant to all simulation functions (this will be indicated in the following chapters). However, there are a set of parameters that apply to all simulations:

+ `verbose [=TRUE]`: Whether extra console output is desired during simulations.
+ `modifyInput [= TRUE]`: Boolean flag to indicate that simulations will modify the input object. If set to `FALSE`, simulations will not modify the input R object but return the current (modified) state variables within the output. In function `fordyn()` option `modifyInput` is always set to FALSE.
+ `fillMissingSpParams [=TRUE]`: Boolean flag to indicate that functions ` spwbInput()` and `growthInput()` should provide estimates for functional parameters if these are lacking in the species parameter table `SpParams`. Note that if `fillMissingSpParams` is set to `FALSE` then simulations may fail if the user does not provide values for required parameters.
+ `subdailyResults [=FALSE]`: Boolean flag to force subdaily results to be stored in calls to `spwb()`. Subdaily results are returned as a list called `subdaily` of `spwb_day` objects, one by simulated date. In function `fordyn()` option `subdailyResults` is always set to `FALSE` and, hence, subdaily results are not possible.

In addition, there is an important control parameter called `transpirationMode`, because it allows the user to switch between the *basic* and *advanced* water balance models (chapters \@ref(basicwaterbalance) and \@ref(advancedwaterbalance)). Another relevant control parameter is `plantWaterPools`, which allows considering plant water pools as explained in \@ref(plantwaterpools).
