# Plant photosynthesis {#plantphotosynthesis}

In the previous chapter we introduced the necessary concepts of steady-state plant hydraulics to define hydraulic supply functions. The supply function specifies the flow rate, as per leaf area, for values of leaf water potential. In this chapter we describe leaf photosynthesis functions, which define leaf gas exchange, photosynthesis and energy balance for values of leaf water potential. In a second section we discuss how to upscale photosynthesis to the crown level.

## Leaf photosynthesis function {#leafenergybalancephoto}

The supply function specifies the flow rate ($E_{leaf}$), as per leaf area, for values of leaf water potential $\Psi_{leaf}$. If we know air temperature, air vapour pressure and the light conditions in which leaves are, we can translate the supply function into a **leaf photosynthesis function** [@Sperry2016]. In a nutshell, $E_{leaf}$ from the supply function is first used to calculate leaf temperature from an evaluation of the leaf energy balance. The diffusive conductances of the leaf to water and $CO_{2}$ are obtained from water supply and water vapor deficit. Finally, the gross/net assimilation rate is then obtained from the diffusive conductance and a modelled curve between assimilation and leaf internal $CO_{2}$ concentration. The following subsections detail each of these steps.


```{r, echo=FALSE}
Tmin = 15
Tmax = 30
RHmin = 60
RHmax = 75
Tcan = meteoland::utils_averageDaylightTemperature(Tmin, Tmax)
vpa = meteoland::utils_averageDailyVP(Tmin, Tmax, RHmin, RHmax)
Patm = meteoland::utils_atmosphericPressure(100)

Q = 2000
Catm = 386

Vmax298 = 100
Jmax298 = 1.67*Vmax298
Gmin = 0.00001;
Gmax = 0.3

Rabs = 740 #W * m-2
```


### Leaf temperature

Leaf temperature ($T_{leaf}$; in Celsius) can be calculated for any given flow rate $E_{leaf}$ using an energy balance equation [@Campbell1998]:
\begin{equation}
T_{leaf} = T_{canopy}+\frac{I_{abs}-\epsilon\cdot\sigma\cdot(T_{canopy}+273.15)^4-\lambda_v\cdot E_{leaf}}{C_p\cdot(g_r+g_{Ha})}
\end{equation}
where $I_{abs}$ (in $W \cdot m^{-2}$) is the instantaneous shortwave and longwave radiation absorbed per leaf area unit, $E_{leaf}$ is the flow (converted to $mol \cdot s^{-1} \cdot m^{-2}$ per two-sided leaf area basis), $\epsilon$ is longwave radiation emissivity (0.97), $\sigma$ is the Stephan-Boltzman constant, $T_{canopy}$ is the canopy air temperature (in ºC) resulting from the energy balance of the previous step (see section \@ref(canopycapacitancetemperature)), $C_p$ = 29.3 $J \cdot mol^{-1} \cdot ºC^{-1}$ is the specific heat capacity of dry air at constant pressure and $\lambda_v$ is the latent heat of vaporization (in $J \cdot mol^{-1}$):
\begin{equation}
\lambda_v = (2.5023\cdot 10^6-(2430.54\cdot T_{canopy}))\cdot 0.018
\end{equation}
Finally, $g_r$ and $g_{Ha}$ are the radiative and heat conductance values (in $mol \cdot m^{-2} \cdot s^{-1}$), respectively [@Campbell1998]:
\begin{eqnarray}
g_r &=& \frac{4\cdot \epsilon \cdot \sigma \cdot (T_{canopy}+273.15)^3}{C_p} \\
g_{Ha} &=& 0.189 \cdot (u_{leaf}/d)^{0.5}
\end{eqnarray}
where $u_{leaf}$ is the leaf-level wind speed (in $m \cdot s^{-1}$), taken as the wind speed at mid-crown height of the plant cohort (see \@ref(windextinctionprofile)), and $d$ is 0.72 times the leaf width (species parameter `LeafWidth` in $cm$). 


Transpiration decreases leaf temperature, whereas radiation increases it and wind speed makes it more similar to air temperature. The following figures illustrate the effect of varying wind speed and flow rate on $T_{leaf}$ for two leaf sizes (see function `biophysics_leafTemperature`):


(ref:leaftemperaturewind-cap) Values of $T_{leaf}$ for two leaf sizes and varying values of wind speed and flow rate, calculated for 24ºC canopy temperature and 740 $W \cdot m^{-2}$ instantaneous absorved radiation. 


```{r leaftemperaturewind, echo=FALSE, fig.width=8, fig.height=4, fig.align="center", fig.cap='(ref:leaftemperaturewind-cap)'}
par(mar=c(4,4,2,1), mfrow=c(1,2))
uvec = seq(0.1, 5, by=0.1)
E = seq(0, 8, by=0.1)
LT1 = matrix(0, nrow= length(uvec), ncol=length(E))
LT2 = matrix(0, nrow= length(uvec), ncol=length(E))
for(i in 1:length(uvec)) {
  for(j in 1:length(E)){
    LT1[i,j] = biophysics_leafTemperature(Rabs,Tcan, uvec[i], E[j], 10.0)
    LT2[i,j] = biophysics_leafTemperature(Rabs,Tcan, uvec[i], E[j], 0.1)
  }
} 
contour(x=uvec, y=E, z=LT1, xlab="Wind speed (m/s)", ylab="Flow rate (mmol·s-1·m-2)", main = "Wide leaf (10 cm) temperature (ºC)")
contour(x=uvec, y=E, z=LT2, xlab="Wind speed (m/s)", ylab="Flow rate (mmol·s-1·m-2)", main = "Narrow leaf (0.1 cm) temperature (ºC)")
```


Let us now fix wind speed at the leaf leavel to $u_{leaf} = 2$ m/s. The application of the above equations to the $E(\Psi_{leaf})$ curves corresponding to the complete hydraulic network (\@ref(supplycontinuum)) yields the following $T_{leaf}(\Psi_{leaf})$ curves:


(ref:leaftemperaturefunction-cap) Examples of leaf temperature functions for a hydraulic network, corresponding to fig. \@ref(fig:supplynetwork) and for different soil textures. Left/right panel shows values for uncavitated/cavitated supply functions. 

```{r leaftemperaturefunction, echo=FALSE, fig.width=8, fig.height=4, fig.align="center", fig.cap='(ref:leaftemperaturefunction-cap)'}
u = 2 # m*s-1
par(mar=c(4,4,2,1), mfrow=c(1,2))
psi2A11 = photo_leafPhotosynthesisFunction(supplyNetwork11$E,Catm, Patm,Tcan, vpa, u, Rabs, Q, Vmax298, Jmax298, Gmin, Gmax)
psi2A12 = photo_leafPhotosynthesisFunction(supplyNetwork12$E,Catm, Patm,Tcan, vpa, u, Rabs, Q, Vmax298, Jmax298, Gmin, Gmax)
psi2A13 = photo_leafPhotosynthesisFunction(supplyNetwork13$E,Catm, Patm,Tcan, vpa, u, Rabs, Q, Vmax298, Jmax298, Gmin, Gmax)
psi2A14 = photo_leafPhotosynthesisFunction(supplyNetwork14$E,Catm, Patm,Tcan, vpa, u, Rabs, Q, Vmax298, Jmax298, Gmin, Gmax)
psi2A21 = photo_leafPhotosynthesisFunction(supplyNetwork21$E,Catm, Patm,Tcan, vpa, u, Rabs, Q, Vmax298, Jmax298, Gmin, Gmax)
psi2A22 = photo_leafPhotosynthesisFunction(supplyNetwork22$E,Catm, Patm,Tcan, vpa, u, Rabs, Q, Vmax298, Jmax298, Gmin, Gmax)
psi2A23 = photo_leafPhotosynthesisFunction(supplyNetwork23$E,Catm, Patm,Tcan, vpa, u, Rabs, Q, Vmax298, Jmax298, Gmin, Gmax)
psi2A24 = photo_leafPhotosynthesisFunction(supplyNetwork24$E,Catm, Patm,Tcan, vpa, u, Rabs, Q, Vmax298, Jmax298, Gmin, Gmax)
psi2A31 = photo_leafPhotosynthesisFunction(supplyNetwork31$E,Catm, Patm,Tcan, vpa, u, Rabs, Q, Vmax298, Jmax298, Gmin, Gmax)
psi2A32 = photo_leafPhotosynthesisFunction(supplyNetwork32$E,Catm, Patm,Tcan, vpa, u, Rabs, Q, Vmax298, Jmax298, Gmin, Gmax)
psi2A33 = photo_leafPhotosynthesisFunction(supplyNetwork33$E,Catm, Patm,Tcan, vpa, u, Rabs, Q, Vmax298, Jmax298, Gmin, Gmax)
psi2A34 = photo_leafPhotosynthesisFunction(supplyNetwork34$E,Catm, Patm,Tcan, vpa, u, Rabs, Q, Vmax298, Jmax298, Gmin, Gmax)

psi2A11emb = photo_leafPhotosynthesisFunction(supplyNetwork11emb$E,Catm, Patm,Tcan, vpa, u, Rabs, Q, Vmax298, Jmax298, Gmin, Gmax)
psi2A12emb = photo_leafPhotosynthesisFunction(supplyNetwork12emb$E,Catm, Patm,Tcan, vpa, u, Rabs, Q, Vmax298, Jmax298, Gmin, Gmax)
psi2A13emb = photo_leafPhotosynthesisFunction(supplyNetwork13emb$E,Catm, Patm,Tcan, vpa, u, Rabs, Q, Vmax298, Jmax298, Gmin, Gmax)
psi2A14emb = photo_leafPhotosynthesisFunction(supplyNetwork14emb$E,Catm, Patm,Tcan, vpa, u, Rabs, Q, Vmax298, Jmax298, Gmin, Gmax)
psi2A21emb = photo_leafPhotosynthesisFunction(supplyNetwork21emb$E,Catm, Patm,Tcan, vpa, u, Rabs, Q, Vmax298, Jmax298, Gmin, Gmax)
psi2A22emb = photo_leafPhotosynthesisFunction(supplyNetwork22emb$E,Catm, Patm,Tcan, vpa, u, Rabs, Q, Vmax298, Jmax298, Gmin, Gmax)
psi2A23emb = photo_leafPhotosynthesisFunction(supplyNetwork23emb$E,Catm, Patm,Tcan, vpa, u, Rabs, Q, Vmax298, Jmax298, Gmin, Gmax)
psi2A24emb = photo_leafPhotosynthesisFunction(supplyNetwork24emb$E,Catm, Patm,Tcan, vpa, u, Rabs, Q, Vmax298, Jmax298, Gmin, Gmax)
psi2A31emb = photo_leafPhotosynthesisFunction(supplyNetwork31emb$E,Catm, Patm,Tcan, vpa, u, Rabs, Q, Vmax298, Jmax298, Gmin, Gmax)
psi2A32emb = photo_leafPhotosynthesisFunction(supplyNetwork32emb$E,Catm, Patm,Tcan, vpa, u, Rabs, Q, Vmax298, Jmax298, Gmin, Gmax)
psi2A33emb = photo_leafPhotosynthesisFunction(supplyNetwork33emb$E,Catm, Patm,Tcan, vpa, u, Rabs, Q, Vmax298, Jmax298, Gmin, Gmax)
psi2A34emb = photo_leafPhotosynthesisFunction(supplyNetwork34emb$E,Catm, Patm,Tcan, vpa, u, Rabs, Q, Vmax298, Jmax298, Gmin, Gmax)

plot(-supplyNetwork11$psiLeaf, psi2A11$LeafTemperature, type="l", col=col1, ylab="Leaf temperature (ºC)", xlab = "Leaf pressure (-MPa)",  xlim=c(0, 7), ylim=c(26,27.5), lwd=1, main="original")
lines(-supplyNetwork12$psiLeaf,psi2A12$LeafTemperature, lty=1, lwd=1, col=col2)
lines(-supplyNetwork13$psiLeaf, psi2A13$LeafTemperature, lty=1, lwd=1, col=col3)
lines(-supplyNetwork14$psiLeaf, psi2A14$LeafTemperature, lty=1, lwd=1, col=col4)
lines(-supplyNetwork21$psiLeaf, psi2A21$LeafTemperature, lty=2, lwd=1, col=col1)
lines(-supplyNetwork22$psiLeaf, psi2A22$LeafTemperature, lty=2, lwd=1, col=col2)
lines(-supplyNetwork23$psiLeaf, psi2A23$LeafTemperature, lty=2, lwd=1, col=col3)
lines(-supplyNetwork24$psiLeaf, psi2A24$LeafTemperature, lty=2, lwd=1, col=col4)
lines(-supplyNetwork31$psiLeaf, psi2A31$LeafTemperature, lty=3, lwd=1, col=col1)
lines(-supplyNetwork32$psiLeaf, psi2A32$LeafTemperature, lty=3, lwd=1, col=col2)
lines(-supplyNetwork33$psiLeaf, psi2A33$LeafTemperature, lty=3, lwd=1, col=col3)
lines(-supplyNetwork34$psiLeaf, psi2A34$LeafTemperature, lty=3, lwd=1, col=col4)
legend("bottomleft", bty="n", legend=textures, lwd=1, lty=1:3, cex=0.8)

plot(-supplyNetwork11emb$psiLeaf, psi2A11emb$LeafTemperature, type="l", col=col1, ylab="Leaf temperature (ºC)", xlab = "Leaf pressure (-MPa)",  xlim=c(0, 7), ylim=c(26,27.5), lwd=1, main="after cavitation")
lines(-supplyNetwork12emb$psiLeaf,psi2A12emb$LeafTemperature, lty=1, lwd=1, col=col2)
lines(-supplyNetwork13emb$psiLeaf, psi2A13emb$LeafTemperature, lty=1, lwd=1, col=col3)
lines(-supplyNetwork14emb$psiLeaf, psi2A14emb$LeafTemperature, lty=1, lwd=1, col=col4)
lines(-supplyNetwork21emb$psiLeaf, psi2A21emb$LeafTemperature, lty=2, lwd=1, col=col1)
lines(-supplyNetwork22emb$psiLeaf, psi2A22emb$LeafTemperature, lty=2, lwd=1, col=col2)
lines(-supplyNetwork23emb$psiLeaf, psi2A23emb$LeafTemperature, lty=2, lwd=1, col=col3)
lines(-supplyNetwork24emb$psiLeaf, psi2A24emb$LeafTemperature, lty=2, lwd=1, col=col4)
lines(-supplyNetwork31emb$psiLeaf, psi2A31emb$LeafTemperature, lty=3, lwd=1, col=col1)
lines(-supplyNetwork32emb$psiLeaf, psi2A32emb$LeafTemperature, lty=3, lwd=1, col=col2)
lines(-supplyNetwork33emb$psiLeaf, psi2A33emb$LeafTemperature, lty=3, lwd=1, col=col3)
lines(-supplyNetwork34emb$psiLeaf, psi2A34emb$LeafTemperature, lty=3, lwd=1, col=col4)
abline(v=-psiCav, col="gray", lwd=1.5)
legend("bottomleft", bty="n", legend=textures, lwd=1, lty=1:3, cex=0.8)
```

### Leaf vapor pressure deficit

Vapor pressure deficit in the leaf ($VPD_{leaf}$, in kPa) is calculated as:
\begin{equation}
VPD_{leaf} = VP(T_{leaf})-vp_{day}
\end{equation}
Where $vp_{day}$ is the average daily vapor pressure and $VP(T)$ is a function giving the saturated vapor pressure for temperature $T$. 

As leaf saturated VP decreases when leaf temperature decreases, increasing transpiration decreases leaf VPD as a result of decreasing leaf temperature. To illustrate this effect, let us assume the following values of relative humidity, yielding an average $vp_{day}$:
```{r}
RHmin = 60
RHmax = 75
vpa = utils_averageDailyVP(Tmin, Tmax, RHmin, RHmax)
vpa
```

The application of the above equation to the $T_{leaf}(\Psi_{leaf})$ curves of fig. \@ref(fig:leaftemperaturefunction) yields the following $VPD_{leaf}(\Psi_{leaf})$ curves:

(ref:leafVPDfunction-cap) Examples of leaf vapour pressure deficit ($VPD_{leaf}$) functions for a hydraulic network, corresponding to fig. \@ref(fig:supplynetwork) and for different soil textures. Left/right panel shows values for uncavitated/cavitated supply functions. 

```{r leafVPDfunction, echo=FALSE, fig.width=8, fig.height=3.5, fig.align="center", fig.cap='(ref:leafVPDfunction-cap)'}
par(mar=c(4,4,2,1), mfrow=c(1,2))
plot(-supplyNetwork11$psiLeaf, psi2A11$LeafVPD, type="l", col=col1, ylab="Leaf VPD (kPa)", xlab = "Leaf pressure (-MPa)",  xlim=c(0, 7), lwd=1, ylim=c(1.4,1.72), main="original")
lines(-supplyNetwork12$psiLeaf,psi2A12$LeafVPD, lty=1, lwd=1, col=col2)
lines(-supplyNetwork13$psiLeaf, psi2A13$LeafVPD, lty=1, lwd=1, col=col3)
lines(-supplyNetwork14$psiLeaf, psi2A14$LeafVPD, lty=1, lwd=1, col=col4)
lines(-supplyNetwork21$psiLeaf, psi2A21$LeafVPD, lty=2, lwd=1, col=col1)
lines(-supplyNetwork22$psiLeaf, psi2A22$LeafVPD, lty=2, lwd=1, col=col2)
lines(-supplyNetwork23$psiLeaf, psi2A23$LeafVPD, lty=2, lwd=1, col=col3)
lines(-supplyNetwork24$psiLeaf, psi2A24$LeafVPD, lty=2, lwd=1, col=col4)
lines(-supplyNetwork31$psiLeaf, psi2A31$LeafVPD, lty=3, lwd=1, col=col1)
lines(-supplyNetwork32$psiLeaf, psi2A32$LeafVPD, lty=3, lwd=1, col=col2)
lines(-supplyNetwork33$psiLeaf, psi2A33$LeafVPD, lty=3, lwd=1, col=col3)
lines(-supplyNetwork34$psiLeaf, psi2A34$LeafVPD, lty=3, lwd=1, col=col4)
legend("bottomleft", bty="n", legend=textures, lwd=1, lty=1:3, cex=0.8)

plot(-supplyNetwork11emb$psiLeaf, psi2A11emb$LeafVPD, type="l", col=col1, ylab="Leaf VPD (kPa)", xlab = "Leaf pressure (-MPa)",  xlim=c(0, 7), lwd=1, ylim=c(1.4,1.72), main="after cavitation")
lines(-supplyNetwork12emb$psiLeaf,psi2A12emb$LeafVPD, lty=1, lwd=1, col=col2)
lines(-supplyNetwork13emb$psiLeaf, psi2A13emb$LeafVPD, lty=1, lwd=1, col=col3)
lines(-supplyNetwork14emb$psiLeaf, psi2A14emb$LeafVPD, lty=1, lwd=1, col=col4)
lines(-supplyNetwork21emb$psiLeaf, psi2A21emb$LeafVPD, lty=2, lwd=1, col=col1)
lines(-supplyNetwork22emb$psiLeaf, psi2A22emb$LeafVPD, lty=2, lwd=1, col=col2)
lines(-supplyNetwork23emb$psiLeaf, psi2A23emb$LeafVPD, lty=2, lwd=1, col=col3)
lines(-supplyNetwork24emb$psiLeaf, psi2A24emb$LeafVPD, lty=2, lwd=1, col=col4)
lines(-supplyNetwork31emb$psiLeaf, psi2A31emb$LeafVPD, lty=3, lwd=1, col=col1)
lines(-supplyNetwork32emb$psiLeaf, psi2A32emb$LeafVPD, lty=3, lwd=1, col=col2)
lines(-supplyNetwork33emb$psiLeaf, psi2A33emb$LeafVPD, lty=3, lwd=1, col=col3)
lines(-supplyNetwork34emb$psiLeaf, psi2A34emb$LeafVPD, lty=3, lwd=1, col=col4)
abline(v=-psiCav, col="gray", lwd=1.5)
legend("bottomleft", bty="n", legend=textures, lwd=1, lty=1:3, cex=0.8)
```


### Leaf gas exchange

Leaf conductance to water vapour ($g_{sw}$; in $mol H_2O \cdot s^{-1} \cdot m^{-2}$) and to carbon dioxide ($g_{sc}$; in mol $CO_{2} \cdot s^{-1} \cdot m^{-2}$) are obtained for each value of $E$ (in $mol \cdot s^{-1} \cdot m^{-2}$) and $VPD_{leaf}$ using:
\begin{eqnarray}
g_{sw} &=& E \cdot \frac{P_{atm}}{VPD_{leaf}}\\
g_{sc} &=& g_{sw}/1.6
\end{eqnarray}
the application of the equation for $g_{sw}$ to the $VPD_{leaf}(\Psi_{leaf})$ curves yields the following $g_{sw}(\Psi_{leaf})$ curves:

```{r, echo=FALSE, fig.width=8, fig.height=3.5, fig.align="center"}
par(mar=c(4,4,2,1), mfrow=c(1,2))
plot(-supplyNetwork11$psiLeaf, psi2A11$WaterVaporConductance, type="l", col=col1, ylab="Gw", xlab = "Leaf pressure (-MPa)",  xlim=c(0, 7), ylim=c(0,0.5), lwd=1, main="original")
lines(-supplyNetwork12$psiLeaf,psi2A12$WaterVaporConductance, lty=1, lwd=1, col=col2)
lines(-supplyNetwork13$psiLeaf, psi2A13$WaterVaporConductance, lty=1, lwd=1, col=col3)
lines(-supplyNetwork14$psiLeaf, psi2A14$WaterVaporConductance, lty=1, lwd=1, col=col4)
lines(-supplyNetwork21$psiLeaf, psi2A21$WaterVaporConductance, lty=2, lwd=1, col=col1)
lines(-supplyNetwork22$psiLeaf, psi2A22$WaterVaporConductance, lty=2, lwd=1, col=col2)
lines(-supplyNetwork23$psiLeaf, psi2A23$WaterVaporConductance, lty=2, lwd=1, col=col3)
lines(-supplyNetwork24$psiLeaf, psi2A24$WaterVaporConductance, lty=2, lwd=1, col=col4)
lines(-supplyNetwork31$psiLeaf, psi2A31$WaterVaporConductance, lty=3, lwd=1, col=col1)
lines(-supplyNetwork32$psiLeaf, psi2A32$WaterVaporConductance, lty=3, lwd=1, col=col2)
lines(-supplyNetwork33$psiLeaf, psi2A33$WaterVaporConductance, lty=3, lwd=1, col=col3)
lines(-supplyNetwork34$psiLeaf, psi2A34$WaterVaporConductance, lty=3, lwd=1, col=col4)
abline(h=Gmin, col="gray", lty=2, lwd=1.5)
abline(h=Gmax, col="gray", lty=2, lwd=1.5)


plot(-supplyNetwork11emb$psiLeaf, psi2A11emb$WaterVaporConductance, type="l", col=col1, ylab="Gw", xlab = "Leaf pressure (-MPa)",  xlim=c(0, 7), ylim=c(0,0.5), lwd=1, main="after cavitation")
lines(-supplyNetwork12emb$psiLeaf,psi2A12emb$WaterVaporConductance, lty=1, lwd=1, col=col2)
lines(-supplyNetwork13emb$psiLeaf, psi2A13emb$WaterVaporConductance, lty=1, lwd=1, col=col3)
lines(-supplyNetwork14emb$psiLeaf, psi2A14emb$WaterVaporConductance, lty=1, lwd=1, col=col4)
lines(-supplyNetwork21emb$psiLeaf, psi2A21emb$WaterVaporConductance, lty=2, lwd=1, col=col1)
lines(-supplyNetwork22emb$psiLeaf, psi2A22emb$WaterVaporConductance, lty=2, lwd=1, col=col2)
lines(-supplyNetwork23emb$psiLeaf, psi2A23emb$WaterVaporConductance, lty=2, lwd=1, col=col3)
lines(-supplyNetwork24emb$psiLeaf, psi2A24emb$WaterVaporConductance, lty=2, lwd=1, col=col4)
lines(-supplyNetwork31emb$psiLeaf, psi2A31emb$WaterVaporConductance, lty=3, lwd=1, col=col1)
lines(-supplyNetwork32emb$psiLeaf, psi2A32emb$WaterVaporConductance, lty=3, lwd=1, col=col2)
lines(-supplyNetwork33emb$psiLeaf, psi2A33emb$WaterVaporConductance, lty=3, lwd=1, col=col3)
lines(-supplyNetwork34emb$psiLeaf, psi2A34emb$WaterVaporConductance, lty=3, lwd=1, col=col4)
abline(v=-psiCav, col="gray", lwd=1.5)
abline(h=Gmin, col="gray", lty=2, lwd=1.5)
abline(h=Gmax, col="gray", lty=2, lwd=1.5)
legend("topleft", bty="n", legend=textures, lwd=1, lty=1:3, cex=0.8)
```

Hence, larger values of transpiration require larger values of leaf water vapour conductance. In the previous figure we have indicated the thresholds of $g_{swmin}$ and $g_{swmax}$, the species-specific minimum and maximum water vapour conductances (i.e. conductances when stomata are fully closed and fully open, respectively; see parameters `Gwmin` and `Gwmax` in `SpParamsMED`).
```{r}
Gmin = 0.0045;
Gmax = 0.3
```

$g_{sw}$ cannot exceed $g_{swmax}$ so that some flow rates may not be possible (see stomatal regulation below). However, $g_{swmax}$ should quickly become non-limiting as soil dries (i.e. reducing $E$) or $VPD_{leaf}$ increases [@Sperry2016]. Minimum stomatal conductance is also used in **medfate** when building the supply function, as it specifies the minimum flow rates that will occur for completely-closed stomata, i.e. the minimum flow from which supply function is build.

### Leaf photosynthesis
Rubisco-limited photosynthesis rate $A_c$ (in $\mu mol CO_2 \cdot s^{-1} \cdot m^{-2}$) is modelled using [@Collatz1991; @Medlyn2002]:
\begin{equation}
A_c=\frac{V_{max}\cdot (C_i- \Gamma*)}{C_i+K_c \cdot (1+ O_a/K_o)}
\end{equation}
where $V_{max}$ is Rubisco's maximum carboxylation rate (in $\mu mol CO_2 \cdot s^{-1} \cdot m^{-2}$), $C_i$ is the internal carbon dioxide concentration (in $\mu mol \cdot mol^{-1}$), $\Gamma*$ is the compensation point (in $\mu mol \cdot mol^{-1}$), $K_c$ (in $\mu mol \cdot mol^{-1}$) and $K_o$ (in $mmol \cdot mol^{-1}$) are Michaelis-Menten constants for carboxylation and oxygenation, respectively, and $O_a$ is the atmospheric oxygen concentration (i.e. 209 $mmol \cdot mol^{-1}$). $\Gamma*$, $K_c$ and $K_o$ depend on leaf temperature ($T_{leaf}$, in Celsius) [@Bernacchi2001]:
\begin{eqnarray}
\Gamma* &=& 42.75\cdot e^{\frac{37830\cdot (T_{leaf}-25)}{298\cdot R \cdot (T_{leaf}-273)}}\\
K_c &=& 404.9\cdot e^{\frac{79430\cdot (T_{leaf}-25)}{298\cdot R \cdot (T_{leaf}-273)}}\\
K_o &=& 278.4\cdot e^{\frac{36380\cdot (T_{leaf}-25)}{298\cdot R \cdot (T_{leaf}-273)}}
\end{eqnarray}
Electron transport-limited photosynthesis $A_e$ (in $\mu mol CO_2 \cdot s^{-1} \cdot m^{-2}$) was obtained from @Medlyn2002:
\begin{eqnarray}
A_e &=& \frac{J}{4}\cdot \frac{C_i-\Gamma*}{C_i+2\cdot \Gamma*} \\
J &=& \frac{(\alpha\cdot Q + J_{max})-\sqrt{(\alpha\cdot Q + J_{max})^2-4.0\cdot c \cdot \alpha \cdot Q \cdot J_{max}}}{2\cdot c}
\end{eqnarray}
where $\alpha$ is the quantum yield of electron transport (0.3 $mol electrons \cdot mol photons^{-1}$), $Q$ is the PAR photon flux density ($\mu mol photons \cdot m^{-2} \cdot s^{-1}$), which is calculated from leaf irradiance ($I_{par}$; in $W \cdot m^{-2}$):
\begin{equation}
Q = I_{par}\cdot 546 \cdot 0.836\cdot 10^{-2}
\end{equation}
$J_{max}$ and $J$ are the maximum and actual rate of electron transport (both in $\mu mol electrons \cdot m^{-2} \cdot s^{-1}$) and $c=0.9$ defines the curvature of the light-response curve. The gross assimilation rate $A$ at a given $C_i$ is the minimum of $A_e$ and $A_c$. To obtain a smooth $A$-vs-$C_i$ curve we used [@Collatz1991]:
\begin{equation}
A = \frac{(A_c+A_e)-\sqrt{(A_c+A_e)^2-4.0\cdot c'\cdot A_e\cdot A_c}}{2\cdot c'}
\end{equation}
where $c'=0.98$ is a curvature factor. The temperature dependence of $J_{max}$ and $V_{max}$ relative to 25ºC was modelled using @Leuning2002 (his eq. 1 with parameters from his Table 2). The internal CO$_2$ concentration, $C_i$, needs to be known to calculate $A$ using the previous equations. @Sperry2016a use a second equation for $A$ which uses $g_{cs}$:
\begin{equation}
A = g_{sc} \cdot (C_{atm}-C_i)
\end{equation}
where $C_{atm}$ is the atmospheric $CO_{2}$ concentration (in $\mu mol \cdot mol^{-1}$; see parameter \texttt{Catm} in function `defaultControl()`). Combining the two equations for $A$ and finding the root of the resulting equation using Newton-Raphson method allows determining $C_i$ and therefore $A$. Thus, after defining PAR photon flux density, atmosphere $CO_{2}$ concentration and maximum rate parameters:
```{r}
Q = 2000
Catm = 386
Vmax298 = 100
Jmax298 = 1.67*Vmax298
```

one can obtain the following $A(\Psi_{leaf})$ curves from $T_{leaf}(\Psi_{leaf})$ and $g_{sc}(\Psi_{leaf})$:

```{r, echo=FALSE, fig.width=8, fig.height=3.5, fig.align="center"}
par(mar=c(4,4,2,1), mfrow=c(1,2))
plot(-supplyNetwork11$psiLeaf, psi2A11$Photosynthesis, type="l", col=col1, ylab="Assimilation rate", xlab = "Leaf sap pressure (-MPa)",  xlim=c(0, 7), lwd=1, main="original")
lines(-supplyNetwork12$psiLeaf,psi2A12$Photosynthesis, lty=1, lwd=1, col=col2)
lines(-supplyNetwork13$psiLeaf, psi2A13$Photosynthesis, lty=1, lwd=1, col=col3)
lines(-supplyNetwork14$psiLeaf, psi2A14$Photosynthesis, lty=1, lwd=1, col=col4)
lines(-supplyNetwork21$psiLeaf, psi2A21$Photosynthesis, lty=2, lwd=1, col=col1)
lines(-supplyNetwork22$psiLeaf, psi2A22$Photosynthesis, lty=2, lwd=1, col=col2)
lines(-supplyNetwork23$psiLeaf, psi2A23$Photosynthesis, lty=2, lwd=1, col=col3)
lines(-supplyNetwork24$psiLeaf, psi2A24$Photosynthesis, lty=2, lwd=1, col=col4)
lines(-supplyNetwork31$psiLeaf, psi2A31$Photosynthesis, lty=3, lwd=1, col=col1)
lines(-supplyNetwork32$psiLeaf, psi2A32$Photosynthesis, lty=3, lwd=1, col=col2)
lines(-supplyNetwork33$psiLeaf, psi2A33$Photosynthesis, lty=3, lwd=1, col=col3)
lines(-supplyNetwork34$psiLeaf, psi2A34$Photosynthesis, lty=3, lwd=1, col=col4)
legend("topright", bty="n", legend=textures, lwd=1, lty=1:3, cex=0.8)

plot(-supplyNetwork11emb$psiLeaf, psi2A11emb$Photosynthesis, type="l", col=col1, ylab="Assimilation rate", xlab = "Leaf sap pressure (-MPa)",  xlim=c(0, 7), lwd=1, main="after cavitation")
lines(-supplyNetwork12emb$psiLeaf,psi2A12emb$Photosynthesis, lty=1, lwd=1, col=col2)
lines(-supplyNetwork13emb$psiLeaf, psi2A13emb$Photosynthesis, lty=1, lwd=1, col=col3)
lines(-supplyNetwork14emb$psiLeaf, psi2A14emb$Photosynthesis, lty=1, lwd=1, col=col4)
lines(-supplyNetwork21emb$psiLeaf, psi2A21emb$Photosynthesis, lty=2, lwd=1, col=col1)
lines(-supplyNetwork22emb$psiLeaf, psi2A22emb$Photosynthesis, lty=2, lwd=1, col=col2)
lines(-supplyNetwork23emb$psiLeaf, psi2A23emb$Photosynthesis, lty=2, lwd=1, col=col3)
lines(-supplyNetwork24emb$psiLeaf, psi2A24emb$Photosynthesis, lty=2, lwd=1, col=col4)
lines(-supplyNetwork31emb$psiLeaf, psi2A31emb$Photosynthesis, lty=3, lwd=1, col=col1)
lines(-supplyNetwork32emb$psiLeaf, psi2A32emb$Photosynthesis, lty=3, lwd=1, col=col2)
lines(-supplyNetwork33emb$psiLeaf, psi2A33emb$Photosynthesis, lty=3, lwd=1, col=col3)
lines(-supplyNetwork34emb$psiLeaf, psi2A34emb$Photosynthesis, lty=3, lwd=1, col=col4)
abline(v=-psiCav, col="gray", lwd=1.5)
legend("topright", bty="n", legend=textures, lwd=1, lty=1:3, cex=0.8)
```

Finally, leaf net photosynthesis (i.e. accounting for autotrophic respiration) is calculated as:
\begin{equation}
A_n = A - 0.015 \cdot V_{max}
\end{equation}


## Crown photosynthesis

In the previous subsection we calculated photosynthesis at the leaf level. However, the function $A(\Psi_{leaf})$ can be calculated for a whole crown. Essentially we need to repeat the calculations of leaf temperature, leaf VPD, leaf gas conductance and photosynthesis for every leaf to be considered in the crown. Gross and net photosynthesis values can be then aggregated across the crown for each value of $\Psi_{leaf}$, so that the function $A(\Psi_{leaf})$ is obtained. Here we will consider a crown of one species divided into 10 layers, with constant leaf density:
```{r}
LAI = 2
nlayer = 10
LAIlayerlive = matrix(rep(LAI/nlayer,nlayer),nlayer,1)
LAIlayermax = matrix(rep(LAI/nlayer,nlayer),nlayer,1)
LAIlayerdead = matrix(0,nlayer,1)
kb = 0.8
kd_PAR = 0.5
kd_SWR = kd_PAR/1.35
alpha_PAR = 0.9
gamma_PAR = 0.04
gamma_SWR = 0.05
alpha_SWR = 0.7
```


Many aspects may vary across the crown, including environmental conditions (such as direct/diffuse light or wind speed) and photosynthesis parameters (e.g. `Vmax298`). The previous crown definition and light parameters lead to a percentage of the above-canopy irradiance reaching each layer [@Anten2016]. Furthermore, it is generally accepted that sunlit and shade leaves need to be treated separately [@DePury1997].Extinction of direct radiation also defines the proportion of leaves of each layer that are affected by direct light beams (i.e. the proportion of sunlit leaves).

```{r, echo=FALSE, fig.width=7, fig.height=4, fig.align="center"}
Ibfpar = light_layerIrradianceFraction(LAIlayerlive,LAIlayerdead,LAIlayermax, kb, alpha_PAR)
Idfpar = light_layerIrradianceFraction(LAIlayerlive,LAIlayerdead,LAIlayermax, kd_PAR, alpha_PAR)
Ibfswr = light_layerIrradianceFraction(LAIlayerlive,LAIlayerdead,LAIlayermax, kb, alpha_SWR)
Idfswr = light_layerIrradianceFraction(LAIlayerlive,LAIlayerdead,LAIlayermax, kd_SWR, alpha_SWR)
fsunlit = light_layerSunlitFraction(LAIlayerlive, LAIlayerdead, kb)
SHarea = (1-fsunlit)*LAIlayerlive[,1] 
SLarea = fsunlit*LAIlayerlive[,1] 

par(mar=c(4,4,1,1), mfrow=c(1,2))
plot(Ibfpar*100, 1:nlayer,type="l", ylab="Layer", xlab="Percentage of irradiance", xlim=c(0,100), ylim=c(1,nlayer), col="dark green")
lines(Idfpar*100, 1:nlayer, col="dark green", lty=2)
lines(Ibfswr*100, 1:nlayer, col="red")
lines(Idfswr*100, 1:nlayer, col="red", lty=2)

plot(fsunlit*100, 1:nlayer,type="l", ylab="Layer", xlab="Percentage of leaves", xlim=c(0,100), ylim=c(1,nlayer))
lines((1-fsunlit)*100, 1:nlayer, lty=2)
```

For simplicity, here we will assume constant windspeed in all layers:
```{r}
ulayer = rep(2, 10)
```

Regarding incoming light, we assume the following direct and diffuse irradiance at the top of the canopy:
```{r}
solarElevation = 0.67
SWR_direct = 1100
SWR_diffuse = 300
PAR_direct = 550
PAR_diffuse = 150
```

Solar elevation is the angle between the sun and the horizon (i.e. the complement of the zenith angle). Under these conditions, the amount of shortwave and PAR radiation absorbed per unit of leaf area at each canopy layer is [@Anten2016]:

```{r, echo=FALSE, fig.width=7, fig.height=4, fig.align="center"}
abs_PAR = light_cohortSunlitShadeAbsorbedRadiation(PAR_direct, PAR_diffuse,
                                             Ibfpar, Idfpar, beta = solarElevation, 
                                             LAIlayerlive, LAIlayerdead, kb, kd_PAR, alpha_PAR, gamma_PAR)
abs_SWR = light_cohortSunlitShadeAbsorbedRadiation(SWR_direct, SWR_diffuse,
                                             Ibfswr, Idfswr, beta = solarElevation, 
                                             LAIlayerlive, LAIlayerdead, kb, kd_SWR, alpha_SWR, gamma_SWR)
par(mar=c(4,4,1,1), mfrow=c(1,2))
absRadSL = abs_SWR$I_sunlit[,1]
absRadSH = abs_SWR$I_shade[,1]
lambda = 546.6507
QSL = abs_PAR$I_sunlit[,1]*lambda*0.836*0.01
QSH = abs_PAR$I_shade[,1]*lambda*0.836*0.01
plot(QSL, 1:nlayer,type="l", ylab="Layer", xlab="Absorbed PAR quantum flux per leaf area", ylim=c(1,nlayer), col="dark green", xlim=c(0,max(QSL)))
lines(QSH, 1:nlayer, col="dark green", lty=2)
plot(absRadSL, 1:nlayer,type="l", ylab="Layer", xlab="Absorbed SWR per leaf area (W/m2)", ylim=c(1,nlayer), col="red", xlim=c(0,max(absRadSL)))
lines(absRadSH, 1:nlayer, col="red", lty=2)
```

Following @DePury1997, we further assume that maximum assimilation rates are highest for leaves at the top of the canopy and there is a exponential decrease from there towards the bottom, where maximum rates are 50% of those at the top: 
\begin{equation}
V_{max,298}(L_i) =V_{max,298}\cdot exp(-0.713\cdot L_i/LAIc)   
\end{equation}
where $L_i$ is the cumulative LAI value at a given canopy layer $i$ and $LAIc$ is the canopy LAI.
```{r, echo=FALSE}
kn = 0.713 #~50% decrease from top to bottom of the canopy
Vmax298layer = numeric(nlayer)
s = 0
for(i in nlayer:1) {
  Vmax298layer[i]=Vmax298*1.5*exp(-kn*(s+LAIlayerlive[i,1]/2)/sum(LAIlayerlive[,1]))
  s =s+LAIlayerlive[i,1]
}
Jmax298layer=Vmax298layer*1.67
# plot(Vmax298layer, 1:nlayer,type="l", ylab="Layer", xlab="Maximum carboxylation rate", ylim=c(1,nlayer), col="red")
```

Multilayer canopy models allow evaluating leaf conditions, stomatal conductance and photosynthesis for different points of the canopy. However, this comes at high computational cost. While big-leaf canopy models are known to be unaccurate under some situations, sun-shade canopy models [@DePury1997] provide estimates that are close to multiple layer models [@Hikosaka2016]. Sun-shade models involve: (a) aggregating the leaf area of sunlit/shade leaves across layers; (b) aggregating the light absorbed by leaves of each kind across layers; and (c) aggregating maximum assimilation rates across layers, again separating sunlit and shade leaves. One then calls the photosynthesis model twice (i.e. once for shade leaves and once for sunlit leaves), using the aggregated maximum assimilation rates. Separating the two kinds of leaves acknowledges that they operate at different parts of the light-saturation curve. The following figure provides the canopy photosynthesis functions obtained, under different situations, using a full 10-layer canopy description (top), a sunshade canopy model (center) or a big-leaf model (bottom). These were generated using functions `photo_multilayerPhotosynthesisFunction()`, `photo_sunshadePhotosynthesisFunction()` and `photo_leafPhotosynthesisFunction()`, respectively. Note the coincidence between the multi-layer and the sun-shade models.
```{r, echo=FALSE}
psi2A11can = photo_multilayerPhotosynthesisFunction(supplyNetwork11$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298layer, Jmax298layer, Gmin, Gmax)
psi2A12can = photo_multilayerPhotosynthesisFunction(supplyNetwork12$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298layer, Jmax298layer, Gmin, Gmax)
psi2A13can = photo_multilayerPhotosynthesisFunction(supplyNetwork13$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298layer, Jmax298layer, Gmin, Gmax)
psi2A14can = photo_multilayerPhotosynthesisFunction(supplyNetwork14$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298layer, Jmax298layer, Gmin, Gmax)
psi2A21can = photo_multilayerPhotosynthesisFunction(supplyNetwork21$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298layer, Jmax298layer, Gmin, Gmax)
psi2A22can = photo_multilayerPhotosynthesisFunction(supplyNetwork22$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298layer, Jmax298layer, Gmin, Gmax)
psi2A23can = photo_multilayerPhotosynthesisFunction(supplyNetwork23$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298layer, Jmax298layer, Gmin, Gmax)
psi2A24can = photo_multilayerPhotosynthesisFunction(supplyNetwork24$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298layer, Jmax298layer, Gmin, Gmax)
psi2A31can = photo_multilayerPhotosynthesisFunction(supplyNetwork31$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298layer, Jmax298layer, Gmin, Gmax)
psi2A32can = photo_multilayerPhotosynthesisFunction(supplyNetwork32$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298layer, Jmax298layer, Gmin, Gmax)
psi2A33can = photo_multilayerPhotosynthesisFunction(supplyNetwork33$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298layer, Jmax298layer, Gmin, Gmax)
psi2A34can = photo_multilayerPhotosynthesisFunction(supplyNetwork34$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298layer, Jmax298layer, Gmin, Gmax)

psi2A11canemb = photo_multilayerPhotosynthesisFunction(supplyNetwork11emb$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298layer, Jmax298layer, Gmin, Gmax)
psi2A12canemb = photo_multilayerPhotosynthesisFunction(supplyNetwork12emb$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298layer, Jmax298layer, Gmin, Gmax)
psi2A13canemb = photo_multilayerPhotosynthesisFunction(supplyNetwork13emb$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298layer, Jmax298layer, Gmin, Gmax)
psi2A14canemb = photo_multilayerPhotosynthesisFunction(supplyNetwork14emb$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298layer, Jmax298layer, Gmin, Gmax)
psi2A21canemb = photo_multilayerPhotosynthesisFunction(supplyNetwork21emb$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298layer, Jmax298layer, Gmin, Gmax)
psi2A22canemb = photo_multilayerPhotosynthesisFunction(supplyNetwork22emb$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298layer, Jmax298layer, Gmin, Gmax)
psi2A23canemb = photo_multilayerPhotosynthesisFunction(supplyNetwork23emb$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298layer, Jmax298layer, Gmin, Gmax)
psi2A24canemb = photo_multilayerPhotosynthesisFunction(supplyNetwork24emb$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298layer, Jmax298layer, Gmin, Gmax)
psi2A31canemb = photo_multilayerPhotosynthesisFunction(supplyNetwork31emb$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298layer, Jmax298layer, Gmin, Gmax)
psi2A32canemb = photo_multilayerPhotosynthesisFunction(supplyNetwork32emb$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298layer, Jmax298layer, Gmin, Gmax)
psi2A33canemb = photo_multilayerPhotosynthesisFunction(supplyNetwork33emb$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298layer, Jmax298layer, Gmin, Gmax)
psi2A34canemb = photo_multilayerPhotosynthesisFunction(supplyNetwork34emb$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298layer, Jmax298layer, Gmin, Gmax)


SLarea = sum(SLarea)
SHarea = sum(SHarea)
QSL = sum(QSL*LAIlayerlive[,1]*fsunlit)
QSH = sum(QSH*LAIlayerlive[,1]*(1-fsunlit))
absRadSL = sum(absRadSL*LAIlayerlive[,1]*fsunlit)
absRadSH = sum(absRadSH*LAIlayerlive[,1]*(1-fsunlit))
Vmax298SL= sum(Vmax298layer*LAIlayerlive[,1]*fsunlit)
Jmax298SL = sum(Jmax298layer*LAIlayerlive[,1]*fsunlit)
Vmax298SH= sum(Vmax298layer*LAIlayerlive[,1]*(1-fsunlit))
Jmax298SH = sum(Jmax298layer*LAIlayerlive[,1]*(1-fsunlit))
ulayer = ulayer[1]

psi2A11ss = photo_sunshadePhotosynthesisFunction(supplyNetwork11$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298SL, Vmax298SH, Jmax298SL, Jmax298SH, Gmin, Gmax)
psi2A12ss = photo_sunshadePhotosynthesisFunction(supplyNetwork12$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298SL, Vmax298SH, Jmax298SL, Jmax298SH, Gmin, Gmax)
psi2A13ss = photo_sunshadePhotosynthesisFunction(supplyNetwork13$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298SL, Vmax298SH, Jmax298SL, Jmax298SH, Gmin, Gmax)
psi2A14ss = photo_sunshadePhotosynthesisFunction(supplyNetwork14$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298SL, Vmax298SH, Jmax298SL, Jmax298SH, Gmin, Gmax)
psi2A21ss = photo_sunshadePhotosynthesisFunction(supplyNetwork21$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298SL, Vmax298SH, Jmax298SL, Jmax298SH, Gmin, Gmax)
psi2A22ss = photo_sunshadePhotosynthesisFunction(supplyNetwork22$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298SL, Vmax298SH, Jmax298SL, Jmax298SH, Gmin, Gmax)
psi2A23ss = photo_sunshadePhotosynthesisFunction(supplyNetwork23$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298SL, Vmax298SH, Jmax298SL, Jmax298SH, Gmin, Gmax)
psi2A24ss = photo_sunshadePhotosynthesisFunction(supplyNetwork24$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298SL, Vmax298SH, Jmax298SL, Jmax298SH, Gmin, Gmax)
psi2A31ss = photo_sunshadePhotosynthesisFunction(supplyNetwork31$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298SL, Vmax298SH, Jmax298SL, Jmax298SH, Gmin, Gmax)
psi2A32ss = photo_sunshadePhotosynthesisFunction(supplyNetwork32$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298SL, Vmax298SH, Jmax298SL, Jmax298SH, Gmin, Gmax)
psi2A33ss = photo_sunshadePhotosynthesisFunction(supplyNetwork33$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298SL, Vmax298SH, Jmax298SL, Jmax298SH, Gmin, Gmax)
psi2A34ss = photo_sunshadePhotosynthesisFunction(supplyNetwork34$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298SL, Vmax298SH, Jmax298SL, Jmax298SH, Gmin, Gmax)
psi2A11ssemb = photo_sunshadePhotosynthesisFunction(supplyNetwork11emb$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298SL, Vmax298SH, Jmax298SL, Jmax298SH, Gmin, Gmax)
psi2A12ssemb = photo_sunshadePhotosynthesisFunction(supplyNetwork12emb$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298SL, Vmax298SH, Jmax298SL, Jmax298SH, Gmin, Gmax)
psi2A13ssemb = photo_sunshadePhotosynthesisFunction(supplyNetwork13emb$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298SL, Vmax298SH, Jmax298SL, Jmax298SH, Gmin, Gmax)
psi2A14ssemb = photo_sunshadePhotosynthesisFunction(supplyNetwork14emb$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298SL, Vmax298SH, Jmax298SL, Jmax298SH, Gmin, Gmax)
psi2A21ssemb = photo_sunshadePhotosynthesisFunction(supplyNetwork21emb$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298SL, Vmax298SH, Jmax298SL, Jmax298SH, Gmin, Gmax)
psi2A22ssemb = photo_sunshadePhotosynthesisFunction(supplyNetwork22emb$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298SL, Vmax298SH, Jmax298SL, Jmax298SH, Gmin, Gmax)
psi2A23ssemb = photo_sunshadePhotosynthesisFunction(supplyNetwork23emb$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298SL, Vmax298SH, Jmax298SL, Jmax298SH, Gmin, Gmax)
psi2A24ssemb = photo_sunshadePhotosynthesisFunction(supplyNetwork24emb$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298SL, Vmax298SH, Jmax298SL, Jmax298SH, Gmin, Gmax)
psi2A31ssemb = photo_sunshadePhotosynthesisFunction(supplyNetwork31emb$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298SL, Vmax298SH, Jmax298SL, Jmax298SH, Gmin, Gmax)
psi2A32ssemb = photo_sunshadePhotosynthesisFunction(supplyNetwork32emb$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298SL, Vmax298SH, Jmax298SL, Jmax298SH, Gmin, Gmax)
psi2A33ssemb = photo_sunshadePhotosynthesisFunction(supplyNetwork33emb$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298SL, Vmax298SH, Jmax298SL, Jmax298SH, Gmin, Gmax)
psi2A34ssemb = photo_sunshadePhotosynthesisFunction(supplyNetwork34emb$E,Catm, Patm,Tcan, vpa, 
                                                SLarea, SHarea, ulayer, 
                                                absRadSL, absRadSH, QSL, QSH,
                                                Vmax298SL, Vmax298SH, Jmax298SL, Jmax298SH, Gmin, Gmax)

BLarea = SLarea+SHarea
QBL = QSL+QSH
absRadBL = absRadSL+absRadSH
Vmax298BL = Vmax298SL + Vmax298SH
Jmax298BL = Jmax298SL + Jmax298SH
Gmax = 2.0
psi2A11bl = photo_leafPhotosynthesisFunction(supplyNetwork11$E,Catm, Patm,Tcan, vpa, ulayer,
                                             absRadBL,QBL,Vmax298BL, Jmax298BL, Gmin, Gmax, 1.0, BLarea)
psi2A12bl = photo_leafPhotosynthesisFunction(supplyNetwork12$E,Catm, Patm,Tcan, vpa,  ulayer,
                                             absRadBL,QBL,Vmax298BL, Jmax298BL, Gmin, Gmax, 1.0, BLarea)
psi2A13bl = photo_leafPhotosynthesisFunction(supplyNetwork13$E,Catm, Patm,Tcan, vpa,  ulayer,
                                             absRadBL,QBL,Vmax298BL, Jmax298BL, Gmin, Gmax, 1.0, BLarea)
psi2A14bl = photo_leafPhotosynthesisFunction(supplyNetwork14$E,Catm, Patm,Tcan, vpa,  ulayer,
                                             absRadBL,QBL,Vmax298BL, Jmax298BL, Gmin, Gmax, 1.0, BLarea)
psi2A21bl = photo_leafPhotosynthesisFunction(supplyNetwork21$E,Catm, Patm,Tcan, vpa,  ulayer,
                                             absRadBL,QBL,Vmax298BL, Jmax298BL, Gmin, Gmax, 1.0, BLarea)
psi2A22bl = photo_leafPhotosynthesisFunction(supplyNetwork22$E,Catm, Patm,Tcan, vpa,  ulayer,
                                             absRadBL,QBL,Vmax298BL, Jmax298BL, Gmin, Gmax, 1.0, BLarea)
psi2A23bl = photo_leafPhotosynthesisFunction(supplyNetwork23$E,Catm, Patm,Tcan, vpa,  ulayer,
                                             absRadBL,QBL,Vmax298BL, Jmax298BL, Gmin, Gmax, 1.0, BLarea)
psi2A24bl = photo_leafPhotosynthesisFunction(supplyNetwork24$E,Catm, Patm,Tcan, vpa,  ulayer,
                                             absRadBL,QBL,Vmax298BL, Jmax298BL, Gmin, Gmax, 1.0, BLarea)
psi2A31bl = photo_leafPhotosynthesisFunction(supplyNetwork31$E,Catm, Patm,Tcan, vpa,  ulayer,
                                             absRadBL,QBL,Vmax298BL, Jmax298BL, Gmin, Gmax, 1.0, BLarea)
psi2A32bl = photo_leafPhotosynthesisFunction(supplyNetwork32$E,Catm, Patm,Tcan, vpa,  ulayer,
                                             absRadBL,QBL,Vmax298BL, Jmax298BL, Gmin, Gmax, 1.0, BLarea)
psi2A33bl = photo_leafPhotosynthesisFunction(supplyNetwork33$E,Catm, Patm,Tcan, vpa,  ulayer,
                                             absRadBL,QBL,Vmax298BL, Jmax298BL, Gmin, Gmax, 1.0, BLarea)
psi2A34bl = photo_leafPhotosynthesisFunction(supplyNetwork34$E,Catm, Patm,Tcan, vpa,  ulayer,
                                             absRadBL,QBL,Vmax298BL, Jmax298BL, Gmin, Gmax, 1.0, BLarea)

psi2A11blemb = photo_leafPhotosynthesisFunction(supplyNetwork11emb$E,Catm, Patm,Tcan, vpa, ulayer,
                                             absRadBL,QBL,Vmax298BL, Jmax298BL, Gmin, Gmax, 1.0, BLarea)
psi2A12blemb = photo_leafPhotosynthesisFunction(supplyNetwork12emb$E,Catm, Patm,Tcan, vpa,  ulayer,
                                             absRadBL,QBL,Vmax298BL, Jmax298BL, Gmin, Gmax, 1.0, BLarea)
psi2A13blemb = photo_leafPhotosynthesisFunction(supplyNetwork13emb$E,Catm, Patm,Tcan, vpa,  ulayer,
                                             absRadBL,QBL,Vmax298BL, Jmax298BL, Gmin, Gmax, 1.0, BLarea)
psi2A14blemb = photo_leafPhotosynthesisFunction(supplyNetwork14emb$E,Catm, Patm,Tcan, vpa,  ulayer,
                                             absRadBL,QBL,Vmax298BL, Jmax298BL, Gmin, Gmax, 1.0, BLarea)
psi2A21blemb = photo_leafPhotosynthesisFunction(supplyNetwork21emb$E,Catm, Patm,Tcan, vpa,  ulayer,
                                             absRadBL,QBL,Vmax298BL, Jmax298BL, Gmin, Gmax, 1.0, BLarea)
psi2A22blemb = photo_leafPhotosynthesisFunction(supplyNetwork22emb$E,Catm, Patm,Tcan, vpa,  ulayer,
                                             absRadBL,QBL,Vmax298BL, Jmax298BL, Gmin, Gmax, 1.0, BLarea)
psi2A23blemb = photo_leafPhotosynthesisFunction(supplyNetwork23emb$E,Catm, Patm,Tcan, vpa,  ulayer,
                                             absRadBL,QBL,Vmax298BL, Jmax298BL, Gmin, Gmax, 1.0, BLarea)
psi2A24blemb = photo_leafPhotosynthesisFunction(supplyNetwork24emb$E,Catm, Patm,Tcan, vpa,  ulayer,
                                             absRadBL,QBL,Vmax298BL, Jmax298BL, Gmin, Gmax, 1.0, BLarea)
psi2A31blemb = photo_leafPhotosynthesisFunction(supplyNetwork31emb$E,Catm, Patm,Tcan, vpa,  ulayer,
                                             absRadBL,QBL,Vmax298BL, Jmax298BL, Gmin, Gmax, 1.0, BLarea)
psi2A32blemb = photo_leafPhotosynthesisFunction(supplyNetwork32emb$E,Catm, Patm,Tcan, vpa,  ulayer,
                                             absRadBL,QBL,Vmax298BL, Jmax298BL, Gmin, Gmax, 1.0, BLarea)
psi2A33blemb = photo_leafPhotosynthesisFunction(supplyNetwork33emb$E,Catm, Patm,Tcan, vpa,  ulayer,
                                             absRadBL,QBL,Vmax298BL, Jmax298BL, Gmin, Gmax, 1.0, BLarea)
psi2A34blemb = photo_leafPhotosynthesisFunction(supplyNetwork34emb$E,Catm, Patm,Tcan, vpa,  ulayer,
                                             absRadBL,QBL,Vmax298BL, Jmax298BL, Gmin, Gmax, 1.0, BLarea)
Gmax = 0.5
```


```{r, echo=FALSE, fig.width=7, fig.height=10, fig.align="center"}
par(mar=c(4,4,2,1), mfrow=c(3,2))
plot(-supplyNetwork11$psiLeaf, psi2A11can$Photosynthesis, type="l", col=col1, ylab="Multilayer photosynthesis", xlab = "Leaf pressure (-MPa)",  xlim=c(0, 7), lwd=1, main="original")
lines(-supplyNetwork12$psiLeaf,psi2A12can$Photosynthesis, lty=1, lwd=1, col=col2)
lines(-supplyNetwork13$psiLeaf, psi2A13can$Photosynthesis, lty=1, lwd=1, col=col3)
lines(-supplyNetwork14$psiLeaf, psi2A14can$Photosynthesis, lty=1, lwd=1, col=col4)
lines(-supplyNetwork21$psiLeaf, psi2A21can$Photosynthesis, lty=2, lwd=1, col=col1)
lines(-supplyNetwork22$psiLeaf, psi2A22can$Photosynthesis, lty=2, lwd=1, col=col2)
lines(-supplyNetwork23$psiLeaf, psi2A23can$Photosynthesis, lty=2, lwd=1, col=col3)
lines(-supplyNetwork24$psiLeaf, psi2A24can$Photosynthesis, lty=2, lwd=1, col=col4)
lines(-supplyNetwork31$psiLeaf, psi2A31can$Photosynthesis, lty=3, lwd=1, col=col1)
lines(-supplyNetwork32$psiLeaf, psi2A32can$Photosynthesis, lty=3, lwd=1, col=col2)
lines(-supplyNetwork33$psiLeaf, psi2A33can$Photosynthesis, lty=3, lwd=1, col=col3)
lines(-supplyNetwork34$psiLeaf, psi2A34can$Photosynthesis, lty=3, lwd=1, col=col4)
legend("bottomright", bty="n", legend=textures, lwd=1, lty=1:3, cex=0.8)

plot(-supplyNetwork11emb$psiLeaf, psi2A11canemb$Photosynthesis, type="l", col=col1, ylab="Multilayer photosynthesis", xlab = "Leaf pressure (-MPa)",  xlim=c(0, 7), lwd=1, main="after cavitation")
lines(-supplyNetwork12emb$psiLeaf,psi2A12canemb$Photosynthesis, lty=1, lwd=1, col=col2)
lines(-supplyNetwork13emb$psiLeaf, psi2A13canemb$Photosynthesis, lty=1, lwd=1, col=col3)
lines(-supplyNetwork14emb$psiLeaf, psi2A14canemb$Photosynthesis, lty=1, lwd=1, col=col4)
lines(-supplyNetwork21emb$psiLeaf, psi2A21canemb$Photosynthesis, lty=2, lwd=1, col=col1)
lines(-supplyNetwork22emb$psiLeaf, psi2A22canemb$Photosynthesis, lty=2, lwd=1, col=col2)
lines(-supplyNetwork23emb$psiLeaf, psi2A23canemb$Photosynthesis, lty=2, lwd=1, col=col3)
lines(-supplyNetwork24emb$psiLeaf, psi2A24canemb$Photosynthesis, lty=2, lwd=1, col=col4)
lines(-supplyNetwork31emb$psiLeaf, psi2A31canemb$Photosynthesis, lty=3, lwd=1, col=col1)
lines(-supplyNetwork32emb$psiLeaf, psi2A32canemb$Photosynthesis, lty=3, lwd=1, col=col2)
lines(-supplyNetwork33emb$psiLeaf, psi2A33canemb$Photosynthesis, lty=3, lwd=1, col=col3)
lines(-supplyNetwork34emb$psiLeaf, psi2A34canemb$Photosynthesis, lty=3, lwd=1, col=col4)
legend("bottomright", bty="n", legend=textures, lwd=1, lty=1:3, cex=0.8)

plot(-supplyNetwork11$psiLeaf, psi2A11ss$Photosynthesis, type="l", col=col1, ylab="Sun-shade photosynthesis", xlab = "Leaf pressure (-MPa)",  xlim=c(0, 7), lwd=1, main="original")
lines(-supplyNetwork12$psiLeaf,psi2A12ss$Photosynthesis, lty=1, lwd=1, col=col2)
lines(-supplyNetwork13$psiLeaf, psi2A13ss$Photosynthesis, lty=1, lwd=1, col=col3)
lines(-supplyNetwork14$psiLeaf, psi2A14ss$Photosynthesis, lty=1, lwd=1, col=col4)
lines(-supplyNetwork21$psiLeaf, psi2A21ss$Photosynthesis, lty=2, lwd=1, col=col1)
lines(-supplyNetwork22$psiLeaf, psi2A22ss$Photosynthesis, lty=2, lwd=1, col=col2)
lines(-supplyNetwork23$psiLeaf, psi2A23ss$Photosynthesis, lty=2, lwd=1, col=col3)
lines(-supplyNetwork24$psiLeaf, psi2A24ss$Photosynthesis, lty=2, lwd=1, col=col4)
lines(-supplyNetwork31$psiLeaf, psi2A31ss$Photosynthesis, lty=3, lwd=1, col=col1)
lines(-supplyNetwork32$psiLeaf, psi2A32ss$Photosynthesis, lty=3, lwd=1, col=col2)
lines(-supplyNetwork33$psiLeaf, psi2A33ss$Photosynthesis, lty=3, lwd=1, col=col3)
lines(-supplyNetwork34$psiLeaf, psi2A34ss$Photosynthesis, lty=3, lwd=1, col=col4)
legend("bottomright", bty="n", legend=textures, lwd=1, lty=1:3, cex=0.8)

plot(-supplyNetwork11emb$psiLeaf, psi2A11ssemb$Photosynthesis, type="l", col=col1, ylab="Sun-shade photosynthesis", xlab = "Leaf pressure (-MPa)",  xlim=c(0, 7), lwd=1, main="after cavitation")
lines(-supplyNetwork12emb$psiLeaf,psi2A12ssemb$Photosynthesis, lty=1, lwd=1, col=col2)
lines(-supplyNetwork13emb$psiLeaf, psi2A13ssemb$Photosynthesis, lty=1, lwd=1, col=col3)
lines(-supplyNetwork14emb$psiLeaf, psi2A14ssemb$Photosynthesis, lty=1, lwd=1, col=col4)
lines(-supplyNetwork21emb$psiLeaf, psi2A21ssemb$Photosynthesis, lty=2, lwd=1, col=col1)
lines(-supplyNetwork22emb$psiLeaf, psi2A22ssemb$Photosynthesis, lty=2, lwd=1, col=col2)
lines(-supplyNetwork23emb$psiLeaf, psi2A23ssemb$Photosynthesis, lty=2, lwd=1, col=col3)
lines(-supplyNetwork24emb$psiLeaf, psi2A24ssemb$Photosynthesis, lty=2, lwd=1, col=col4)
lines(-supplyNetwork31emb$psiLeaf, psi2A31ssemb$Photosynthesis, lty=3, lwd=1, col=col1)
lines(-supplyNetwork32emb$psiLeaf, psi2A32ssemb$Photosynthesis, lty=3, lwd=1, col=col2)
lines(-supplyNetwork33emb$psiLeaf, psi2A33ssemb$Photosynthesis, lty=3, lwd=1, col=col3)
lines(-supplyNetwork34emb$psiLeaf, psi2A34ssemb$Photosynthesis, lty=3, lwd=1, col=col4)
legend("bottomright", bty="n", legend=textures, lwd=1, lty=1:3, cex=0.8)

plot(-supplyNetwork11$psiLeaf, psi2A11bl$Photosynthesis, type="l", col=col1, ylab="Big-leaf photosynthesis", xlab = "Leaf pressure (-MPa)",  xlim=c(0, 7), lwd=1, main="original")
lines(-supplyNetwork12$psiLeaf,psi2A12bl$Photosynthesis, lty=1, lwd=1, col=col2)
lines(-supplyNetwork13$psiLeaf, psi2A13bl$Photosynthesis, lty=1, lwd=1, col=col3)
lines(-supplyNetwork14$psiLeaf, psi2A14bl$Photosynthesis, lty=1, lwd=1, col=col4)
lines(-supplyNetwork21$psiLeaf, psi2A21bl$Photosynthesis, lty=2, lwd=1, col=col1)
lines(-supplyNetwork22$psiLeaf, psi2A22bl$Photosynthesis, lty=2, lwd=1, col=col2)
lines(-supplyNetwork23$psiLeaf, psi2A23bl$Photosynthesis, lty=2, lwd=1, col=col3)
lines(-supplyNetwork24$psiLeaf, psi2A24bl$Photosynthesis, lty=2, lwd=1, col=col4)
lines(-supplyNetwork31$psiLeaf, psi2A31bl$Photosynthesis, lty=3, lwd=1, col=col1)
lines(-supplyNetwork32$psiLeaf, psi2A32bl$Photosynthesis, lty=3, lwd=1, col=col2)
lines(-supplyNetwork33$psiLeaf, psi2A33bl$Photosynthesis, lty=3, lwd=1, col=col3)
lines(-supplyNetwork34$psiLeaf, psi2A34bl$Photosynthesis, lty=3, lwd=1, col=col4)
legend("bottomright", bty="n", legend=textures, lwd=1, lty=1:3, cex=0.8)

plot(-supplyNetwork11emb$psiLeaf, psi2A11blemb$Photosynthesis, type="l", col=col1, ylab="Big-leaf photosynthesis", xlab = "Leaf pressure (-MPa)",  xlim=c(0, 7), lwd=1, main="after cavitation")
lines(-supplyNetwork12emb$psiLeaf,psi2A12blemb$Photosynthesis, lty=1, lwd=1, col=col2)
lines(-supplyNetwork13emb$psiLeaf, psi2A13blemb$Photosynthesis, lty=1, lwd=1, col=col3)
lines(-supplyNetwork14emb$psiLeaf, psi2A14blemb$Photosynthesis, lty=1, lwd=1, col=col4)
lines(-supplyNetwork21emb$psiLeaf, psi2A21blemb$Photosynthesis, lty=2, lwd=1, col=col1)
lines(-supplyNetwork22emb$psiLeaf, psi2A22blemb$Photosynthesis, lty=2, lwd=1, col=col2)
lines(-supplyNetwork23emb$psiLeaf, psi2A23blemb$Photosynthesis, lty=2, lwd=1, col=col3)
lines(-supplyNetwork24emb$psiLeaf, psi2A24blemb$Photosynthesis, lty=2, lwd=1, col=col4)
lines(-supplyNetwork31emb$psiLeaf, psi2A31blemb$Photosynthesis, lty=3, lwd=1, col=col1)
lines(-supplyNetwork32emb$psiLeaf, psi2A32blemb$Photosynthesis, lty=3, lwd=1, col=col2)
lines(-supplyNetwork33emb$psiLeaf, psi2A33blemb$Photosynthesis, lty=3, lwd=1, col=col3)
lines(-supplyNetwork34emb$psiLeaf, psi2A34blemb$Photosynthesis, lty=3, lwd=1, col=col4)
legend("bottomright", bty="n", legend=textures, lwd=1, lty=1:3, cex=0.8)
```


