# Decomposition and soil biogeochemistry {#decomposition}
TO BE WRITTEN

\begin{equation}
k_{i} = k_{\text{base,}i} \cdot f(T) \cdot f(\theta) \cdot f(pH)
\end{equation}

Where $k_{\text{base,}i}$ is the base daily decomposition rate for pool $i$ which is multiplied by three modifiers: $ƒ(T)$ is the effect of temperature, $f(\theta_{rel})$ is the effect of moisture relative to saturation, and $f(pH)$ is the effect of soil pH.

### Effect of pH

The effect of pH on decomposition, $f(pH)$ is calculated using:

\begin{equation}
f(pH) = b + \frac{c}{\pi} \cdot \text{tan}^{-1} [d \cdot (pH - a) \cdot \pi]
\end{equation}

where parameters $a$, $b$, $c$ and $d$ depend on the specific carbon pool:

| Carbon pool(s)          | $a$ | $b$ | $c$ | $d$ |
|-------------------------|-----|-----|-----|-----|
| Surface/soil metabolic  | 4.8 | 0.5 | 1.14 | 0.7 |
| Surface/soil structural  | 4.0 | 0.5 | 1.1 | 0.7 |
| Surface active  | 4.0 | 0.5 | 1.1 | 0.7 |
| Soil active  | 4.8 | 0.5 | 1.14 | 0.7 |
| Surface/soil slow  | 4.0 | 0.5 | 1.1 | 0.7 |
| Soil passive  | 3.0 | 0.5 | 1.1 | 0.7 |

(ref:pH-decomp-cap) Effect of pH on decomposition rates, for different carbon pools.

```{r, fig.align='center', fig.width=6, fig.height=3.5, echo = FALSE, fig.cap='(ref:pH-decomp-cap)'}
df1 <- data.frame(pH = seq(4,8, by = 0.001), f = NA, Pool = "Surface/soil metabolic/active")
df2 <- data.frame(pH = seq(4,8, by = 0.001), f = NA, Pool = "Surface/soil structural/slow")
df3 <- data.frame(pH = seq(4,8, by = 0.001), f = NA, Pool = "Soil passive")
for(i in 1:nrow(df1)) {
  df1$f[i] = decomposition_pHEffect(df1$pH[i], "SurfaceMetabolic")
  df2$f[i] = decomposition_pHEffect(df2$pH[i], "Leaves")
  df3$f[i] = decomposition_pHEffect(df3$pH[i], "SoilPassive")
}
df <- rbind(df1, df2, df3)
ggplot(df)+
  geom_line(aes(x = pH, y = f, col = Pool)) +
  scale_color_discrete(name = NULL)+
  ylim(0,1)+
  xlab("pH") + ylab("pH effect on decomposition")+
  theme_bw()+
  theme(legend.position = c(0.8,0.2))
```

### Effect of temperature

The effect of temperature on decomposition, $f(T)$, follows the same general structure as the effect of pH, but with fixed coefficients (i.e. $a = 15.7$, $b = 0.56$, $c = 1.46$ and $d = 0.031$) [@del_grosso_modeling_2005]:

\begin{equation}
f(T) = 0.56 + \frac{1.46}{\pi} \cdot \text{tan}^{-1} [0.0309 \cdot \pi \cdot (T - 15.7) ]
\end{equation}

(ref:temp-decomp-cap) Effect of temperature on decomposition rates.

```{r, fig.align='center', fig.width=6, fig.height=3.5, echo = FALSE, fig.cap='(ref:temp-decomp-cap)'}
df <- data.frame(temperature = seq(0,30, by = 0.1), f = NA)
for(i in 1:nrow(df)) {
  df$f[i] = decomposition_temperatureEffect(df$temperature[i])
}
ggplot(df)+
  geom_line(aes(x = temperature, y = f)) +
  xlab("Temperature in Celsius") + ylab("Temperature effect on decomposition")+
  theme_bw()
```

### Effect of moisture

The effect of moisture is modelled using [@kelly_intra-annual_2000]:

\begin{equation}
f(\theta_{rel}) = \left(\frac{\theta_{rel} - b}{a - b} \right) ^{ \frac{d\cdot(b-a)}{a-c}} \cdot \left( \frac{\theta_{rel} - c}{a - c}\right) ^{d}
\end{equation}

where $\theta_{rel}$ is moisture relative to saturation, and here parameters $a$, $b$, $c$ and $d$ depend on soil texture [@kelly_intra-annual_2000]: 

| Soil texture          | $a$ | $b$ | $c$ | $d$ |
|-------------------------|-----|-----|-----|-----|
| Fine texture  | 0.6 | 1.27 | 0.0012 | 2.84 |
| Coarse texture  | 0.55 | 1.7 | -0.007 | 3.22 |

(ref:moisture-decomp-cap) Effect of moisture relative to saturation ($\theta_{rel}$) on decomposition rates, for fine textured and coarse textured soil.

```{r, fig.align='center', fig.width=6, fig.height=3.5, echo = FALSE, fig.cap='(ref:moisture-decomp-cap)'}
df1 <- data.frame(theta_rel = seq(0.002,0.999, by = 0.001), f = NA, texture = "fine")
df2 <- data.frame(theta_rel = seq(0.002,0.999, by = 0.001), f = NA, texture = "coarse")
for(i in 1:nrow(df1)) {
  df1$f[i] = decomposition_moistureEffect(10,30, df1$theta_rel[i])
  df2$f[i] = decomposition_moistureEffect(30,10, df2$theta_rel[i])
}
df <- rbind(df1, df2)
ggplot(df)+
  geom_line(aes(x = theta_rel, y = f, col = texture)) +
  scale_color_discrete(name = NULL)+
  ylim(0,1)+
  xlab("Relative moisture") + ylab("Moisture effect on decomposition")+
  theme_bw() +
  theme(legend.position = c(0.2,0.8))
```


## Decomposition of standing dead wood and litter elements
TO BE WRITTEN

### Standing dead wood

Two types carbon pools are considered for standing dead wood (fig. \@ref(fig:decompositionpools)): **small branches** (diameters between 0.635 cm and 7.5 cm; 10h and 100h fuels; in $g C \cdot m^{-2}$) and **large wood** (diameters larger than 7.5 cm; 1000h and 10000h fuels; in $g C \cdot m^{-2}$).

The rate of decomposition of a standing dead pool $i$ is:

\begin{equation}
k_{i} = k_{\text{base,}i} \cdot f(T) \cdot f(\theta_{rel}) \cdot \exp(-3 \cdot f_{lig})
\end{equation}

where $k_{\text{base,}i}$ is the base daily decomposition rate (which is higher for small branches than large wood); $f_{lig}$ is the fraction of wood lignin; and atmosphere (or canopy in the case of advanced simulations) air temperature is used for temperature effects  (i.e. $T = T_{air}$ or $T = T_{can}$). Moisture relative to saturation ($\theta_{rel}$) is estimated as the ratio of actual fuel moisture to its maximum value (both with units of percent of water mass to dry mass):

\begin{equation}
\theta_{rel} = \theta_{Simard}(T_{air}, RH_{air})/\theta_{max}
\end{equation}

where $\theta_{Simard}$ is the equilibrium moisture content (Simard 1968) of wood, according to an empirical relationship with air temperature and air relative humidity [@viney_review_1991] and $\theta_{max}$ is the maximum fuel moisture content, which depends on wood density (assuming the density of water as $1\,\, g\cdot cm^{-3}$):

\begin{equation}
\theta_{max} = 100 \cdot \left(\frac{1}{\rho_{wood}} - \frac{1}{\rho_{ws}} \right)
\end{equation}

where $\rho_{wood}$ is wood density ($g\cdot cm^{-3}$) and $\rho_{ws} = 1.53\,\,g\cdot cm^{-3}$ is the density of the wood substance (a constant). Since $\theta_{Simard}$ ranges between 0% and 30%, whereas $\theta_{max}$ ranges between 45% ($\rho_{wood} = 0.9$) and 135% ($\rho_{wood} = 0.5$), the $\theta_{rel}$ ratio is frequently lower than the same ratio assessed from soil moisture, resulting in slower decomposition rates.

### Litter structural and metabolic pools

## Dynamics of soil carbon pools

TO BE WRITTEN
